---
title: "Initial Descriptives"
author: "Group 3: Institutional Characteristics"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: cayman
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
pacman::p_load(here, rio, tidyverse, janitor)
```

## Introduction

We are creating descriptives for the data here.

Research Question:

### Creating one dataset

```{r individual data files}
finance_data <- import(here("Data", "Data Institutional Finance Variables.csv"))  # Dong

dependentdv_data <- import(here("Data", "Data Dependent Variables.csv")) # Dong

studentbody_data <- import(here("Data", "Data Student Body Composition Variables.csv"))  # Anwesha

carnegie_data <- import(here("Data/Data Carnegie Variables", "Data Carnegie Variables.csv")) # Anwesha
#  janitor::clean_names() # omitting this to add in labels first

academic_data <- import(here("Data", "Data Academic Variables.csv")) # Amleset

admissions_data <- import(here("Data", "Data Admissions Variables.csv")) # Amleset

geographic_data <- import(here("Data/Data Geographic Variables", "Data Geographic Variables.csv")) # Kurt

# non-IPEDS data TBD # Kurt
```

The code below begins to clean the finance data and put it in long format.
```{r}
# aimed format: institution, year, variable, value

#starting off with finance data
finance_data <- finance_data %>%
  janitor::clean_names() %>% 
  select(1:2, sort(names(.)[-(1:2)]), -v218) #sort alphabetical to better see vars. also, v218 is an empty variable tacked onto the end, so I removed it here

# most follow the 4-digit year format, except the avg net price i'll remove those, clean them up, and then add them back in.
finance_netprice_clean <- finance_data %>%
  select(unit_id, institution_name, starts_with("average_net_price")) %>% 
  mutate(avg_net_price_w_aid_2019 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2018_19_sfa1819_rv,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2018_19_sfa1819_rv_2),
         avg_net_price_w_aid_2020 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2019_20_sfa1920_rv,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2019_20_sfa1920_rv_2),
         avg_net_price_w_aid_2021 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2020_21_sfa2021_rv,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2020_21_sfa2021_rv_2),
         avg_net_price_w_aid_2022 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2021_22_sfa2122_rv,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2021_22_sfa2122_rv_2),
         avg_net_price_w_aid_2023 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2022_23_sfa2223,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2022_23_sfa2223_2)) %>% 
  select(1, 2, 13:17)

finance_data <- finance_data %>% 
  select(-starts_with("average_net_price"), -starts_with("imputation")) %>%  #removing imputation flag for now
  left_join(., finance_netprice_clean) 

finance_data_clean <- finance_data %>% 
  pivot_longer(
    cols = -c(unit_id, institution_name),
    names_to = "variable",
    values_to = "value"
  ) %>% 
  mutate(
    year = str_extract(variable, "\\d{4}")
  )

## clean up variable name
finance_data_clean <- finance_data_clean %>% 
  mutate(variable_base = str_extract(variable, ".*(?=_(drv|ic))")) %>% #take out the suffix
  select(-variable)
```

On second thought, it might make more sense for everyone to do this part however they like, and then I can take the datasets created and pull them all together in the end. I'll still leave the code here in case it's useful, but I'll move on. - Anwesha

### Student Body & Carnegie Data (Anwesha)

#### Carnegie classification
```{r}
carnegie_labels <- import(here("Data/Data Carnegie Variables", "Labels Carnegie Variables.csv"))

carnegie_data <- carnegie_data %>% 
  select(-V9) %>% # get rid of rogue empty variable
  pivot_longer(
    cols = -c(1:2),
    names_to = "VariableName",
    values_to = "Value"
  ) 

# add in labels
carnegie_labelled <- carnegie_data %>% 
  left_join(., carnegie_labels) %>% 
  select(-Value)
```

Descriptives by Carnegie classification are below. Note that there are about 8 schools that enroll graduate students exclusively. We may want to exclude these schools when we run our analyses.
```{r}
carnegie_labelled <- carnegie_labelled %>% 
  pivot_wider(
    names_from = VariableName,
    values_from = ValueLabel
  )
```

```{r carnegieplots, fig.width=10, fig.height=8}
# Get the names of the variables to plot (excluding UnitID and Institution Name)
vars_to_plot <- names(carnegie_labelled)[3:8]

# Loop over each variable and create a bar chart
for (var in vars_to_plot) {
  
  p <- carnegie_labelled %>%
    count(.data[[var]]) %>%
    ggplot(aes(x = reorder(.data[[var]], n), y = n)) +
    geom_bar(stat = "identity", fill = "#005288") + # color taken from website
    geom_text(aes(label = n), hjust = -0.5, size = 3) +
    labs(title = var, x = NULL, y = "Count") +
    theme_minimal() +
    theme(panel.grid.major.y = element_blank(),
          plot.title = element_text(face = "bold", size = 13, hjust = 0, margin = margin(l = -80))) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.07))) + # bars flush with axes, some space for text
    coord_flip()
  
  print(p)
}
```

#### Student Body

```{r}
studentbody_data <- studentbody_data %>% 
  select(-V88, -contains("Imputation")) %>% 
  pivot_longer(
    cols = -c(1:2),
    names_to = "VariableName",
    values_to = "Value"
  )

studentbody_data_clean <- studentbody_data %>% 
  rename_with(~ str_trim(.), everything()) %>%  # trim any weird spacing in names
  mutate(
    year = case_when(
      str_detect(VariableName, "DRVEF\\d{4}") ~ str_extract(VariableName, "DRVEF(\\d{4})") %>% str_remove("DRVEF"),
      str_detect(VariableName, "SFA\\d{4}") ~ paste0("20", str_sub(str_extract(VariableName, "SFA\\d{4}"), -2)),  # grab last 2 digits
      TRUE ~ NA_character_
    ),  # extract 4-digit year conditionally
    clean_label = str_remove(VariableName, "\\s*\\([^\\)]+\\)")  # remove everything in parentheses
  ) %>% #looks good! I'll remove original variable name
  select(-VariableName)
```

Let's see what the variable names are here.
```{r}
table(studentbody_data_clean$clean_label)
```

The Pell and Non-resident variables are broken into two different variables due to different naming conventions. I'll combine those.
```{r}
studentbody_data_clean <- studentbody_data_clean %>% 
  mutate(clean_label = recode(clean_label, "Percent of undergraduate enrollment that are Nonresident Alien" = 
        "Percent of undergraduate enrollment that are U.S. Nonresident",
        "Percent of undergraduate students awarded Pell grants" = 
        "Percent of undergraduate students awarded Federal Pell grants"))

table(studentbody_data_clean$clean_label)
```
Great, everything is uniform now. I'll pivot the dataset to better create a descriptive table. Note that the schools and/or years with 0 enrollment should be removed from the dataset when we do the analysis.

```{r}
overall_minmax_students <- studentbody_data_clean %>%
  group_by(clean_label) %>%
  summarise(
    n = sum(!is.na(Value)),
    min = min(Value, na.rm = TRUE),
    max = max(Value, na.rm = TRUE),
    .groups = "drop"
  )

studentbody_descriptives <- studentbody_data_clean %>%
  group_by(clean_label, year) %>%
  summarise(
    mean = round(mean(Value, na.rm = TRUE), 1),
    median = round(median(Value, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  # Create year-specific columns
  pivot_wider(
    names_from = year,
    values_from = c(mean, median),
    names_glue = "{.value}_{year}"
  ) %>%
  arrange(clean_label) %>% 
  right_join(overall_minmax_students, by = "clean_label")
```
```{r}
library(kableExtra)

studentbody_descriptives %>%
  kable(digits = 1, 
        col.names = gsub("_", " ", names(studentbody_descriptives)),
        caption = "Summary Statistics by Year") %>%
  kable_styling(full_width = FALSE, 
                bootstrap_options = c("striped", "hover", "condensed"))
```



### Other Variables ... 

Info.

## Exploratory Questions

### Overall Trends

Guiding Questions:

- What is the percent of [institutions] that have seen an increase in undergraduate enrollment [in recent history]? What is the percent of [institutions] that have seen a decrease in undergraduate enrollment [in recent history]? 
- What percent of [institutions] have seen an increase in undergraduate enrollment [in recent history] of [scale, e.g., 10% or more]? What is the percent of [institutions] that have seen a decrease in undergraduate enrollment [in recent history] of [scale, e.g., 10% or more]? 

### Changing Enrollment

Guiding Questions:

- Is change in undergraduate enrollment [in recent history] correlated with change in application volume (number of applicants)? At what scale? [Same year] 
- Is change in undergraduate enrollment [in recent history] correlated with change in yield rate (% of admitted students who enroll at the institution)? At what scale? 
- Is change in undergraduate enrollment [in recent history] correlated with change in “attractor score” (number of applicants * yield rate)? At what scale? 
- Is change in undergraduate enrollment [in recent history] correlated with change in new student enrollment (first-time, first-year and transfer students, together and/or separately)? 

### Student Composition

Which student populations are increasing/decreasing in enrollment? 
- Enrollment by race/ethnicity 
- Percent Pell 
- Adult age (25-64) enrollment 

### Intersection between Enrollment and Graduation

TBD.

