---
title: "Initial Descriptives"
author: "Group 3: Institutional Characteristics"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: cayman
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
pacman::p_load(here, rio, tidyverse, janitor)
```

## Introduction

We are creating descriptives for the data here.

Research Question:

### Creating one dataset

```{r individual data files}
finance_data <- import(here("Data", "Data Institutional Finance Variables.csv"))  # Dong

dependentdv_data <- import(here("Data", "Data Dependent Variables.csv")) # Dong

studentbody_data <- import(here("Data", "Data Student Body Composition Variables.csv"))  # Anwesha

carnegie_data <- import(here("Data/Data Carnegie Variables", "Data Carnegie Variables.csv")) # Anwesha
#  janitor::clean_names() # omitting this to add in labels first

academic_data <- import(here("Data", "Data Academic Variables.csv")) # Amleset

admissions_data <- import(here("Data", "Data Admissions Variables.csv")) # Amleset

# geographic_data <- import(here("Data/Data Geographic Variables", "Data Geographic Variables.csv")) # Kurt
# IPEDS: Geographic data ----
ipeds_geog_values <- read_csv(
  file = "Data/Data Geographic Variables/Data Geographic Variables.csv", col_select = 1:9
)
ipeds_geog_labels <- read_csv(
  file = "Data/Data Geographic Variables/Labels Geographic Variables.csv"
)


# non-IPEDS data TBD # Kurt
# WICHE Data: High School Graduates by Year by State ----
wiche_data_allrows <- readxl::read_xlsx(
  path = "Data/Non-IPEDS Data/WICHE HS Graduates by State.xlsx",
  sheet = "Data"
)
# ACS Data: Educational Attainment by State ----
acs_2018 <- read_csv(
  file = "Data/Non-IPEDS Data/ACS Educational Attainment by State/ACSST1Y2018.S1501-2025-04-09T015452.csv"
)
stabbr_csv <- read_csv(
  file = "Data/Non-IPEDS Data/states.csv"
)
```

The code below begins to clean the finance data and put it in long format.
```{r}
# aimed format: institution, year, variable, value

#starting off with finance data
finance_data <- finance_data %>%
  janitor::clean_names() %>% 
  select(1:2, sort(names(.)[-(1:2)]), -v218) #sort alphabetical to better see vars. also, v218 is an empty variable tacked onto the end, so I removed it here

# most follow the 4-digit year format, except the avg net price i'll remove those, clean them up, and then add them back in.
finance_netprice_clean <- finance_data %>%
  select(unit_id, institution_name, starts_with("average_net_price")) %>% 
  mutate(avg_net_price_w_aid_2019 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2018_19_sfa1819_rv,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2018_19_sfa1819_rv_2),
         avg_net_price_w_aid_2020 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2019_20_sfa1920_rv,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2019_20_sfa1920_rv_2),
         avg_net_price_w_aid_2021 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2020_21_sfa2021_rv,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2020_21_sfa2021_rv_2),
         avg_net_price_w_aid_2022 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2021_22_sfa2122_rv,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2021_22_sfa2122_rv_2),
         avg_net_price_w_aid_2023 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2022_23_sfa2223,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2022_23_sfa2223_2)) %>% 
  select(1, 2, 13:17)

finance_data <- finance_data %>% 
  select(-starts_with("average_net_price"), -starts_with("imputation")) %>%  #removing imputation flag for now
  left_join(., finance_netprice_clean) 

finance_data_clean <- finance_data %>% 
  pivot_longer(
    cols = -c(unit_id, institution_name),
    names_to = "variable",
    values_to = "value"
  ) %>% 
  mutate(
    year = str_extract(variable, "\\d{4}")
  )

## clean up variable name
#finance_data_clean <- finance_data_clean %>% 
#  mutate(variable_base = str_extract(variable, ".*(?=_(drv|ic))")) %>% #take out the suffix
#  select(-variable)
```
On second thought, it might make more sense for everyone to do this part however they like, and then I can take the datasets created and pull them all together in the end. I'll still leave the code here in case it's useful, but I'll move on. - Anwesha

### Finance data & dependent dv_data (Dong)
#### Finance data
```{r}
## clean up variable name
library(stringi)

finance_data_clean <- finance_data_clean %>%
  mutate(
    variable_base = variable %>%
      str_remove("\\s*\\([^\\)]+\\)") %>%
      str_trim() %>%
      str_to_lower() %>%
      str_replace_all("[^a-z0-9]", "_") %>%
      str_replace_all("_+", "_") %>%
      str_remove("^_") %>%
      str_remove("_$") %>%
      str_remove("(_\\d{4}_\\d{2}_ic\\d{4}_ay)$") %>%       
      str_remove("(_\\d{4}_ic\\d{4}_ay)$") %>%               
      str_remove("(_\\d{4})$") %>%                          
      str_remove("_drv[f]?\\d{4}(_rv)?$") %>%
      str_remove("_sfa\\d{4}(_rv)?$") %>%
      str_remove("_fgr\\d{4}(_rv)?$") %>%
      stri_trans_general("Latin-ASCII")
  ) %>%
  filter(!is.na(variable_base), !is.na(value), !is.na(year))
```

```{r}
# Finance data analysis - select key variables to focus 
finance_data_clean %>%
  distinct(variable_base) %>%
  arrange(variable_base)

finance_keywords <- c(
  "tuition", "net_price", "instruction_expenses", "core_revenues", "core_expenses",
  "appropriations", "student_service_expenses", "academic_support_expenses",
  "research_expenses", "institutional_support_expenses", "investment_return",
  "published", "private_gifts", "government_grants", "public_service_expenses"
)

finance_focus <- finance_data_clean %>%
  filter(str_detect(variable_base, str_c(finance_keywords, collapse = "|")),
         year %in% 2018:2023) %>%
  mutate(value = as.numeric(value))

finance_descriptives_long <- finance_focus %>%
  group_by(variable_base, year) %>%
  summarise(mean = round(mean(value, na.rm = TRUE), 1),
            median = round(median(value, na.rm = TRUE), 1),
            min = round(min(value, na.rm = TRUE), 1),
            max = round(max(value, na.rm = TRUE), 1),
            n = sum(!is.na(value)),
            .groups = "drop") %>%
  pivot_wider(names_from = year,
              values_from = c(mean, median),
              names_glue = "{.value}_{year}")

```

```{r}
# create readable variable names
variable_labels <- tibble::tibble(
  variable_base = c(
    "academic_support_expenses_as_a_percent_of_total_core_expenses_fasb",
    "academic_support_expenses_as_a_percent_of_total_core_expenses_gasb",
    "instruction_expenses_as_a_percent_of_total_core_expenses_fasb",
    "instruction_expenses_as_a_percent_of_total_core_expenses_gasb",
    "institutional_support_expenses_as_a_percent_of_total_core_expenses_fasb",
    "institutional_support_expenses_as_a_percent_of_total_core_expenses_gasb",
    "student_service_expenses_as_a_percent_of_total_core_expenses_fasb",
    "student_service_expenses_as_a_percent_of_total_core_expenses_gasb",
    "tuition_and_fees_as_a_percent_of_core_revenues_fasb",
    "tuition_and_fees_as_a_percent_of_core_revenues_gasb",
    "core_expenses_total_dollars_fasb",
    "core_expenses_total_dollars_gasb",
    "core_revenues_total_dollars_fasb",
    "core_revenues_total_dollars_gasb",
    "average_net_price_students_awarded_grant_or_scholarship_aid_2018_19_sfa1819_rv",
    "average_net_price_students_awarded_grant_or_scholarship_aid_2019_20_sfa1920_rv",
    "average_net_price_students_awarded_grant_or_scholarship_aid_2020_21_sfa2021_rv",
    "average_net_price_students_awarded_grant_or_scholarship_aid_2021_22_sfa2122_rv",
    "average_net_price_students_awarded_grant_or_scholarship_aid_2022_23_sfa2223",
    "avg_net_price_w_aid",
    "government_grants_and_contracts_as_a_percent_of_core_revenues_fasb",
    "government_grants_and_contracts_as_a_percent_of_core_revenues_gasb",
    "instruction_expenses_per_fte_fasb",
    "instruction_expenses_per_fte_gasb",
    "investment_return_as_a_percent_of_core_revenues_fasb",
    "investment_return_as_a_percent_of_core_revenues_gasb",
    "local_appropriations_as_a_percent_of_core_revenues_gasb",
    "other_core_expenses_as_a_percent_of_total_core_expenses_fasb",
    "other_core_expenses_as_a_percent_of_total_core_expenses_gasb",
    "other_revenues_as_a_percent_of_core_revenues_fasb",
    "other_revenues_as_a_percent_of_core_revenues_gasb",
    "private_gifts_grants_and_contracts_as_a_percent_of_core_revenues_gasb",
    "private_gifts_grants_contracts_contributions_from_affiliated_entities_as_a_percent_of_core_revenues_fasb",
    "public_service_expenses_as_a_percent_of_total_core_expenses_fasb",
    "public_service_expenses_as_a_percent_of_total_core_expenses_gasb",
    "published_in_district_tuition_and_fees",
    "published_in_state_tuition_and_fees",
    "published_out_of_state_tuition_and_fees",
    "research_expenses_as_a_percent_of_total_core_expenses_fasb",
    "research_expenses_as_a_percent_of_total_core_expenses_gasb",
    "state_appropriations_as_percent_of_core_revenues_gasb"
  ),
  variable_label = c(
    "Academic Support (% Core Expenses) – FASB",
    "Academic Support (% Core Expenses) – GASB",
    "Instruction (% Core Expenses) – FASB",
    "Instruction (% Core Expenses) – GASB",
    "Institutional Support (% Core Expenses) – FASB",
    "Institutional Support (% Core Expenses) – GASB",
    "Student Services (% Core Expenses) – FASB",
    "Student Services (% Core Expenses) – GASB",
    "Tuition & Fees (% Core Revenues) – FASB",
    "Tuition & Fees (% Core Revenues) – GASB",
    "Total Core Expenses (Dollars) – FASB",
    "Total Core Expenses (Dollars) – GASB",
    "Total Core Revenues (Dollars) – FASB",
    "Total Core Revenues (Dollars) – GASB",
    "Avg Net Price (With Aid) – 2018–19",
    "Avg Net Price (With Aid) – 2019–20",
    "Avg Net Price (With Aid) – 2020–21",
    "Avg Net Price (With Aid) – 2021–22",
    "Avg Net Price (With Aid) – 2022–23",
    "Avg Net Price W Aid",
    "Government Grants And Contracts (% Core Revenues) – FASB",
    "Government Grants And Contracts (% Core Revenues) – GASB",
    "Instruction Expenses Per FTE – FASB",
    "Instruction Expenses Per FTE – GASB",
    "Investment Return (% Core Revenues) – FASB",
    "Investment Return (% Core Revenues) – GASB",
    "Local Appropriations (% Core Revenues) – GASB",
    "Other Core Expenses (% Core Expenses) – FASB",
    "Other Core Expenses (% Core Expenses) – GASB",
    "Other Revenues (% Core Revenues) – FASB",
    "Other Revenues (% Core Revenues) – GASB",
    "Private Gifts Grants And Contracts (% Core Revenues) – GASB",
    "Private Gifts Grants Contributions From Affiliated Entities (% Core Revenues) – FASB",
    "Public Service Expenses (% Core Expenses) – FASB",
    "Public Service Expenses (% Core Expenses) – GASB",
    "Published In-District Tuition And Fees",
    "Published In-State Tuition And Fees",
    "Published Out-Of-State Tuition And Fees",
    "Research Expenses (% Core Expenses) – FASB",
    "Research Expenses (% Core Expenses) – GASB",
    "State Appropriations (% Core Revenues) – GASB"
  )
)

```

```{r}
# clean labels 
variable_labels <- variable_labels %>%
  mutate(
    variable_base = variable_base %>%
      str_remove("\\s*\\([^\\)]+\\)") %>%
      str_trim() %>%
      str_to_lower() %>%
      str_replace_all("[^a-z0-9]", "_") %>%
      str_replace_all("_+", "_") %>%
      str_remove("^_") %>%
      str_remove("_$") %>%
      str_remove("_drvf?\\d{4}_?rv?$") %>%
      str_remove("_sfa\\d{4}_?rv?$") %>%
      str_remove("_fgr\\d{4}_?rv?$") %>%
      stri_trans_general("Latin-ASCII")
  )
```

```{r}
# Merge human-readable variable_label into the main summary table (To make the table easier to understand for readers)
finance_descriptives_labeled <- finance_descriptives_long %>%
  left_join(variable_labels, by = "variable_base") %>%
  relocate(variable_label, .after = variable_base)

finance_summary_merged <- finance_descriptives_labeled %>%
  group_by(variable_base, variable_label) %>%   
  summarise(
    min = min(min, na.rm = TRUE),
    max = max(max, na.rm = TRUE),
    n = sum(n, na.rm = TRUE),
    across(starts_with("mean_"), ~ first(na.omit(.x))),
    across(starts_with("median_"), ~ first(na.omit(.x))),
    .groups = "drop"
  ) %>%
  arrange(variable_label)
```

```{r}
#  Make a descriptive table
finance_summary_merged %>%
  kable(
    digits = 1,
    caption = "Finance Variables: Year 2019-2023 Summary"
  ) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"), 
                position = "left") %>%
  scroll_box(width = "100%", height = "500px")
```

#### Dependent dv_data
```{r}
# Data Cleaning and Preparation：Clean column names, reshape to long format, and extract variable base and year
dependent_data_clean <- dependentdv_data %>%
  janitor::clean_names() %>%
  select(where(is.numeric), institution_name) %>%
  pivot_longer(
    cols = -institution_name,
    names_to = "variable",
    values_to = "value"
  ) %>%
  mutate(
    # 1. Extract the 4-digit year (most robust)
    year = str_extract(variable, "20[1-2][0-9]"),  
    
    # 2. Remove year-related substrings to get variable_base
    variable_base = variable %>%
      str_remove_all("20[1-2][0-9]") %>%                          
       str_remove_all("(_)?(19|20|21|22|23)(_|$)") %>%                                      
      str_remove_all("_rv|_drv|_drvef|_gr|_sfa|_fgr") %>%         
      str_remove_all("_august_31|_first_time|_full_time|_enter") %>%   
      str_replace_all("_+", "_") %>%
      str_remove("^_") %>%
      str_remove("_$") %>%
      stri_trans_general("Latin-ASCII")   # Normalize any special characters
  ) %>%
  filter(
    !is.na(variable_base),
    !is.na(value),
    !is.na(year),
    year %in% c("2019", "2020", "2021", "2022", "2023") 
  )

```

```{r}
#Identify and filter key dependent variables by keyword
key_dv_keywords <- c(
  "graduation_rate", "completers_within", "adjusted_cohort",
  "receiving_a_bachelor_s_degree", "undergraduate_enrollment", 
  "full_time_equivalent_enrollment"
)

dependent_keyvars <- dependent_data_clean %>%
  filter(str_detect(variable_base, str_c(key_dv_keywords, collapse = "|")))
```

```{r}
# Generate yearly descriptive statistics for each key variable
dependent_summary <- dependent_keyvars %>%
  mutate(value = as.numeric(value)) %>%
  group_by(variable_base, year) %>%
  summarise(
    n = sum(!is.na(value)),
    min = round(min(value, na.rm = TRUE), 1),
    max = round(max(value, na.rm = TRUE), 1),
    mean = round(mean(value, na.rm = TRUE), 1),
    median = round(median(value, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  arrange(variable_base, year)


# Generate readable labels
dependent_labels <- dependent_summary %>%
  distinct(variable_base) %>%
  mutate(
    variable_label = variable_base %>%
      str_replace_all("_", " ") %>%
      str_to_title() %>%
      str_replace("Bachelor S", "Bachelor's") %>%
      str_replace("Drv", "Graduation Rate") %>%
      str_replace("Drvef", "Enrollment") %>%
      str_replace("Om", "Outcome Measure") %>%
      str_replace("Graduation Rate Total Cohortgr", "Graduation Rate Total Cohort Graduation Rate") %>%   
      str_replace_all("_enrollmentef", "Fall Enrollment")  
  )
```

```{r}
# Merge labels with descriptive statistics and reshape to wide format
dependent_summary_labeled <- dependent_summary %>%
  left_join(dependent_labels, by = "variable_base") %>%
  relocate(variable_label, .after = variable_base)

dependent_summary_wide <- dependent_summary_labeled %>%
  select(variable_base, variable_label, year, mean, median) %>%
  pivot_wider(
    names_from = year,
    values_from = c(mean, median),
    names_glue = "{.value}_{year}"
  ) %>%
  arrange(variable_label)

# Output as formatted table
dependent_summary_wide %>%
  kable(
    digits = 1,
    caption = "Key Dependent Variables Yearly Summary",
    escape = FALSE
  ) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))
```

### Student Body & Carnegie Data (Anwesha)

#### Carnegie classification
```{r}
carnegie_labels <- import(here("Data/Data Carnegie Variables", "Labels Carnegie Variables.csv"))

carnegie_data <- carnegie_data %>% 
  select(-V9) %>% # get rid of rogue empty variable
  pivot_longer(
    cols = -c(1:2),
    names_to = "VariableName",
    values_to = "Value"
  ) 

# add in labels
carnegie_labelled <- carnegie_data %>% 
  left_join(., carnegie_labels) %>% 
  select(-Value)
```

Descriptives by Carnegie classification are below. Note that there are about 8 schools that enroll graduate students exclusively. We may want to exclude these schools when we run our analyses.
```{r}
carnegie_labelled <- carnegie_labelled %>% 
  pivot_wider(
    names_from = VariableName,
    values_from = ValueLabel
  )
```

```{r carnegieplots, fig.width=10, fig.height=8}
# Get the names of the variables to plot (excluding UnitID and Institution Name)
vars_to_plot <- names(carnegie_labelled)[3:8]

# Loop over each variable and create a bar chart
for (var in vars_to_plot) {
  
  p <- carnegie_labelled %>%
    count(.data[[var]]) %>%
    ggplot(aes(x = reorder(.data[[var]], n), y = n)) +
    geom_bar(stat = "identity", fill = "#005288") + # color taken from website
    geom_text(aes(label = n), hjust = -0.5, size = 3) +
    labs(title = var, x = NULL, y = "Count") +
    theme_minimal() +
    theme(panel.grid.major.y = element_blank(),
          plot.title = element_text(face = "bold", size = 13, hjust = 0, margin = margin(l = -80))) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.07))) + # bars flush with axes, some space for text
    coord_flip()
  
  print(p)
}
```

#### Student Body

```{r}
studentbody_data <- studentbody_data %>% 
  select(-V88, -contains("Imputation")) %>% 
  pivot_longer(
    cols = -c(1:2),
    names_to = "VariableName",
    values_to = "Value"
  )

studentbody_data_clean <- studentbody_data %>% 
  rename_with(~ str_trim(.), everything()) %>%  # trim any weird spacing in names
  mutate(
    year = case_when(
      str_detect(VariableName, "DRVEF\\d{4}") ~ str_extract(VariableName, "DRVEF(\\d{4})") %>% str_remove("DRVEF"),
      str_detect(VariableName, "SFA\\d{4}") ~ paste0("20", str_sub(str_extract(VariableName, "SFA\\d{4}"), -2)),  # grab last 2 digits
      TRUE ~ NA_character_
    ),  # extract 4-digit year conditionally
    clean_label = str_remove(VariableName, "\\s*\\([^\\)]+\\)")  # remove everything in parentheses
  ) %>% #looks good! I'll remove original variable name
  select(-VariableName)
```

Let's see what the variable names are here.
```{r}
table(studentbody_data_clean$clean_label)
```

The Pell and Non-resident variables are broken into two different variables due to different naming conventions. I'll combine those.
```{r}
studentbody_data_clean <- studentbody_data_clean %>% 
  mutate(clean_label = recode(clean_label, "Percent of undergraduate enrollment that are Nonresident Alien" = 
        "Percent of undergraduate enrollment that are U.S. Nonresident",
        "Percent of undergraduate students awarded Pell grants" = 
        "Percent of undergraduate students awarded Federal Pell grants"))

table(studentbody_data_clean$clean_label)
```
Great, everything is uniform now. I'll pivot the dataset to better create a descriptive table. Note that the schools and/or years with 0 enrollment should be removed from the dataset when we do the analysis.

```{r}
overall_minmax_students <- studentbody_data_clean %>%
  group_by(clean_label) %>%
  summarise(
    n = sum(!is.na(Value)),
    min = min(Value, na.rm = TRUE),
    max = max(Value, na.rm = TRUE),
    .groups = "drop"
  )

studentbody_descriptives <- studentbody_data_clean %>%
  group_by(clean_label, year) %>%
  summarise(
    mean = round(mean(Value, na.rm = TRUE), 1),
    median = round(median(Value, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  # Create year-specific columns
  pivot_wider(
    names_from = year,
    values_from = c(mean, median),
    names_glue = "{.value}_{year}"
  ) %>%
  arrange(clean_label) %>% 
  right_join(overall_minmax_students, by = "clean_label")
```
```{r}
library(kableExtra)

studentbody_descriptives %>%
  kable(digits = 1, 
        col.names = gsub("_", " ", names(studentbody_descriptives)),
        caption = "Summary Statistics by Year") %>%
  kable_styling(full_width = FALSE, 
                bootstrap_options = c("striped", "hover", "condensed"))
```

### Geographic and non-IPEDS Data (Kurt)

```{r}
# High school graduates by year by state
hs_grads_by_year <- wiche_data_allrows %>%
  filter(
    Grade == "High school graduates", # Keep only rows ~ total high school graduates
    RaceEthnicity == "Total/any", # Keep only rows ~ total all race/ethnicity
    SchoolSector == "Grand Total (public+private)", # Keep only rows ~ total public and private high schools
    !(str_detect(string = Stabbr, pattern = "^_")), # Keep only rows NOT starting with an underscore (i.e., drop region aggregates)
    !(str_detect(string = Stabbr, pattern = "\\d")), # Keep only rows NOT containing digits (i.e., drop region aggregates)
    !(Stabbr %in% c("CM","CW")) # Keep only rows NOT CM and CW (i.e., drop region aggregates)
  ) %>%
  mutate(
    HS_Grad_Year = as.numeric(str_extract(string = ClassOf, pattern = "\\d+")) # Extract the graduation year for numeric filtering
  ) %>%
  filter(
    HS_Grad_Year >= 2017,
    HS_Grad_Year < 2026
  ) %>%
  select(
    # Keep only these variables
    Stabbr,
    HS_Grad_Year,
    Students
  ) %>%
  # Convert to wide format...
  pivot_wider(id_cols = Stabbr, names_from = HS_Grad_Year, values_from = Students, names_prefix = "HS_Grad_")

# Educational attainment (percent adults 25+ with a Bachelors degree or higher) by state 2018
educ_attain_2018 <- select(acs_2018, `Label (Grouping)`, ends_with("Percent!!Estimate"))[16,] %>%
  pivot_longer(cols = 2:53, names_to = "ACS_variable") %>%
  mutate(
    State = str_extract(string = ACS_variable, pattern = "^[^!]+(?=!)")
  ) %>%
  left_join(
    y = stabbr_csv,
    by = "State"
  ) %>%
  rename(ACS_Pct_Bachelors = value) %>%
  mutate(ACS_Pct_Bachelors = as.numeric(str_extract(ACS_Pct_Bachelors, pattern="\\d+\\.\\d+")))

# The analysis variables for the geographic data are the following:
# Bureau of Economic Analysis (BEA) regions (HD2023)
# Degree of urbanization (Urban-centric locale) (HD2023)
# Keep state abbreviation `State abbreviation (HD2023)` for joining the other two tables

ipeds_geog_values$BEA_Region <- factor(
  x = ipeds_geog_values$`Bureau of Economic Analysis (BEA) regions (HD2023)`,
  levels = filter(ipeds_geog_labels, VariableName=="Bureau of Economic Analysis (BEA) regions (HD2023)")$Value %>% as.numeric(),
  labels = filter(ipeds_geog_labels, VariableName=="Bureau of Economic Analysis (BEA) regions (HD2023)")$ValueLabel
)

ipeds_geog_values$Locale <- factor(
  x = ipeds_geog_values$`Degree of urbanization (Urban-centric locale) (HD2023)`,
  levels = filter(ipeds_geog_labels, VariableName=="Degree of urbanization (Urban-centric locale) (HD2023)")$Value %>% as.numeric(),
  labels = filter(ipeds_geog_labels, VariableName=="Degree of urbanization (Urban-centric locale) (HD2023)")$ValueLabel
)

ipeds_geog_values$Locale_Rural <- case_when(
  ipeds_geog_values$`Degree of urbanization (Urban-centric locale) (HD2023)` %in% c(11,12,13,21,22,23) ~ "City/Suburb",
  ipeds_geog_values$`Degree of urbanization (Urban-centric locale) (HD2023)` %in% c(31,32,33,41,42,43) ~ "Town/Rural",
  TRUE ~ NA
)

# Joining geographic tables together ----

ipeds_geog_data <- ipeds_geog_values %>%
  select(
    UnitID,
    `Institution Name`,
    `State abbreviation (HD2023)`,
    BEA_Region,
    Locale,
    Locale_Rural
  ) %>%
  left_join(
    y = hs_grads_by_year,
    by = c("State abbreviation (HD2023)"="Stabbr")
  ) %>%
  left_join(
    y = educ_attain_2018 %>% select(Abbreviation, ACS_Pct_Bachelors),
    by = c("State abbreviation (HD2023)"="Abbreviation")
  )
```

Calculate descriptive statistics for Region, Locale, HS grads by state, and Percent Bachelors holders by state

```{r}
# Calculate descriptive statistics ----
map(
  .x = ipeds_geog_data %>% select(BEA_Region, Locale, HS_Grad_2017, ACS_Pct_Bachelors),
  .f = function(x) {
    if(is.factor(x)) {
      x_tbl <- table(x, useNA = "ifany")
      
      tibble(
        Values = names(x_tbl),
        Val_N = x_tbl,
        Val_Pct = 100*round(x_tbl / sum(x_tbl), 3)
      ) %>%
        arrange(desc(Val_Pct))
    } else if(is.numeric(x)) {
      list(
        fivenum(x),
        c(mean(x), sd(x))
      )
    }
  }
)
```


### Other Variables ... 

Info.

## Exploratory Questions

### Overall Trends

Guiding Questions:

- What is the percent of [institutions] that have seen an increase in undergraduate enrollment [in recent history]? What is the percent of [institutions] that have seen a decrease in undergraduate enrollment [in recent history]? 
- What percent of [institutions] have seen an increase in undergraduate enrollment [in recent history] of [scale, e.g., 10% or more]? What is the percent of [institutions] that have seen a decrease in undergraduate enrollment [in recent history] of [scale, e.g., 10% or more]? 

### Changing Enrollment

Guiding Questions:

- Is change in undergraduate enrollment [in recent history] correlated with change in application volume (number of applicants)? At what scale? [Same year] 
- Is change in undergraduate enrollment [in recent history] correlated with change in yield rate (% of admitted students who enroll at the institution)? At what scale? 
- Is change in undergraduate enrollment [in recent history] correlated with change in “attractor score” (number of applicants * yield rate)? At what scale? 
- Is change in undergraduate enrollment [in recent history] correlated with change in new student enrollment (first-time, first-year and transfer students, together and/or separately)? 

### Student Composition

Which student populations are increasing/decreasing in enrollment? 
- Enrollment by race/ethnicity 
- Percent Pell 
- Adult age (25-64) enrollment 

### Intersection between Enrollment and Graduation

TBD.

