---
title: "Descriptives - Version 2"
author: "Anwesha Guha"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: cayman
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(here, rio, tidyverse, janitor, gt, psych)
```

## Read in data

```{r individual data files, eval=FALSE, include=FALSE}
finance_data <- import(here("Data", "Data Institutional Finance Variables.csv"))  # Dong

dependentdv_data <- import(here("Data", "Data Dependent Variables.csv")) # Dong

studentbody_data <- import(here("Data", "Data Student Body Composition Variables.csv"))  # Anwesha

carnegie_data <- import(here("Data/Data Carnegie Variables", "Data Carnegie Variables.csv")) # Anwesha
#  janitor::clean_names() # omitting this to add in labels first

academic_data <- import(here("Data", "Data Academic Variables.csv")) # Amleset, NOT USING

admissions_data <- import(here("Data", "Data Admissions Variables.csv")) # Amleset

# geographic_data <- import(here("Data/Data Geographic Variables", "Data Geographic Variables.csv")) # Kurt
# IPEDS: Geographic data ----
ipeds_geog_values <- read_csv(
  file = "Data/Data Geographic Variables/Data Geographic Variables.csv", col_select = 1:9
)
ipeds_geog_labels <- read_csv(
  file = "Data/Data Geographic Variables/Labels Geographic Variables.csv"
)


# non-IPEDS data TBD # Kurt
# WICHE Data: High School Graduates by Year by State ----
wiche_data_allrows <- readxl::read_xlsx(
  path = "Data/Non-IPEDS Data/WICHE HS Graduates by State.xlsx",
  sheet = "Data"
)
# ACS Data: Educational Attainment by State ----
acs_2018 <- read_csv(
  file = "Data/Non-IPEDS Data/ACS Educational Attainment by State/ACSST1Y2018.S1501-2025-04-09T015452.csv"
)
stabbr_csv <- read_csv(
  file = "Data/Non-IPEDS Data/states.csv"
)
```

## Dependent variables 

  Enrollment – defined as first-time full-time student cohort 

## Independent variables 


###  Applicants (numeric: integers) 

```{r}
admissions_data <- import(here("Data", "Data Admissions Variables.csv"))

# 1: Select only total count columns (exclude imputation flags)
admission_data_clean <- admissions_data %>%
  clean_names() %>% 
  select(unit_id, institution_name, matches("^applicants|^admissions|^enrolled")) %>%
  pivot_longer(
    cols = -c(unit_id, institution_name),
    names_to = "variable",
    values_to = "value"
  ) %>%
  
  # 2: Extract category and year from column names
  mutate(
    category = case_when(
      str_detect(variable, "applicants") ~ "applicants",
      str_detect(variable, "admissions") ~ "admissions",
      str_detect(variable, "enrolled") ~ "enrolled"
    ),
    year = str_extract(variable, "20\\d{2}"),
    year = as.integer(year)
  ) %>%
  filter(!is.na(year)) %>%  # Remove NA years (some might not match pattern)
  filter(value > 1) # Should have over 1 student enrolled

summary_stats <- admission_data_clean %>%
  group_by(year, category) %>%
  summarise(
    Mean = round(mean(value, na.rm = TRUE)),
    Median = round(median(value, na.rm = TRUE)),
    SD = round(sd(value, na.rm = TRUE), 2),
    Count = n()
  )
```

###  Admissions rate (numeric: %, 0 to 100) 

```{r}
admissions_rate_df <- admission_data_clean %>%
  group_by(unit_id, institution_name, year, category) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = category, values_from = value) %>%
  mutate(admissions_rate = admissions / applicants)
```

```{r}
describe(admissions_rate_df)
```


###  Average net price with aid (numeric: integers) 
```{r}
finance_data <- import(here("Data", "Data Institutional Finance Variables.csv")) %>% 
  janitor::clean_names() %>% 
  select(unit_id, institution_name, starts_with("average_net_price")) %>%
  mutate(avg_net_price_w_aid_2019 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2018_19_sfa1819_rv,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2018_19_sfa1819_rv_2),
         avg_net_price_w_aid_2020 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2019_20_sfa1920_rv,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2019_20_sfa1920_rv_2),
         avg_net_price_w_aid_2021 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2020_21_sfa2021_rv,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2020_21_sfa2021_rv_2),
         avg_net_price_w_aid_2022 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2021_22_sfa2122_rv,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2021_22_sfa2122_rv_2),
         avg_net_price_w_aid_2023 = coalesce(average_net_price_students_awarded_grant_or_scholarship_aid_2022_23_sfa2223,
                                             average_net_price_students_awarded_grant_or_scholarship_aid_2022_23_sfa2223_2)) %>% 
  select(1, 2, 13:17)

finance_data_clean <- finance_data %>%
  pivot_longer(
    cols = starts_with("avg_net_price_w_aid"),  # Select columns with "avg_net_price_w_aid"
    names_to = "year",                         # The new column to hold the year
    values_to = "avg_net_price_w_aid"          # The new column to hold the values
  ) %>%
  mutate(year = as.integer(sub("avg_net_price_w_aid_", "", year))) %>%  # Clean up the 'year' column to integer
  filter(avg_net_price_w_aid > 0) #omit weird negative

```

```{r}
describe(finance_data_clean$avg_net_price_w_aid)
```

```{r}
describeBy(finance_data_clean$avg_net_price_w_aid, group = finance_data_clean$year) #2021 was cheaper
```

###  Percent of undergraduate students awarded Federal Pell grants (numeric: %, 0 to 100) 

```{r}
studentbody_data <- import(here("Data", "Data Student Body Composition Variables.csv"))

studentbody_data <- studentbody_data %>% 
  select(-V88, -contains("Imputation")) %>% 
  pivot_longer(
    cols = -c(1:2),
    names_to = "VariableName",
    values_to = "Value"
  )

studentbody_data_clean <- studentbody_data %>% 
  rename_with(~ str_trim(.), everything()) %>%  # trim any weird spacing in names
  mutate(
    year = case_when(
      str_detect(VariableName, "DRVEF\\d{4}") ~ str_extract(VariableName, "DRVEF(\\d{4})") %>% str_remove("DRVEF"),
      str_detect(VariableName, "SFA\\d{4}") ~ paste0("20", str_sub(str_extract(VariableName, "SFA\\d{4}"), -2)),  # grab last 2 digits
      TRUE ~ NA_character_
    ),  # extract 4-digit year conditionally
    clean_label = str_remove(VariableName, "\\s*\\([^\\)]+\\)")  # remove everything in parentheses
  ) %>% #looks good! I'll remove original variable name
  select(-VariableName)

# The Pell and Non-resident variables are broken into two different variables due to different naming conventions. I'll combine those.

studentbody_data_clean <- studentbody_data_clean %>% 
  mutate(clean_label = recode(clean_label, "Percent of undergraduate enrollment that are Nonresident Alien" = 
        "Percent of undergraduate enrollment that are U.S. Nonresident",
        "Percent of undergraduate students awarded Pell grants" = 
        "Percent of undergraduate students awarded Federal Pell grants"))

# table(studentbody_data_clean$clean_label) #check

minoritized_dat <- studentbody_data_clean %>% 
  mutate(clean_label = case_when(clean_label == "Percent of undergraduate students awarded Federal Pell grants" ~ "pct_pell",
                               clean_label == "Percent of undergraduate enrollment that are White" ~ "pct_white",
                               clean_label == "Undergraduate enrollment" ~ "enrollment")) %>% 
  filter(clean_label == "pct_pell" | clean_label == "pct_white" | clean_label == "enrollment") %>% 
  janitor::clean_names() %>% 
  pivot_wider(names_from = clean_label,
              values_from = value) %>% 
  mutate(pct_race_minority = 100 - pct_white) %>% 
  select(-pct_white) %>% 
  filter(enrollment > 1)
```

###  Percent racially minoritized students (numeric: %, 0 to 100) 

```{r}
describe(minoritized_dat)
```

## Complete dataset

We will restrict the sample to the schools that have all the independent variables represented.

```{r}
minoritized_dat$year = as.integer(minoritized_dat$year)
finance_data_clean$year = as.integer(finance_data_clean$year)
admissions_rate_df$year = as.integer(admissions_rate_df$year)

admissions_selected <- admissions_rate_df %>% select(-enrolled, -admissions)

dat <- inner_join(minoritized_dat, finance_data_clean)
dat <- inner_join(dat, admissions_selected)
```

```{r}
length(unique(dat$unit_id))
```

We have 1,606 unique Institution IDs in our sample.

```{r}
describe(dat) %>% 
  gt() 
#  fmt_number(decimals = 2)
```
Examine by year.

```{r}
dat %>%
  group_by(year) %>%
  summarise(
    across(
      -c(unit_id, institution_name),
      list(mean = mean, sd = sd),
      na.rm = TRUE
    ),
    .groups = "drop"
  ) %>% 
  pivot_longer(
    cols = -year,
    names_to = c("variable", "stat"),
    names_pattern = "^(.*)_(mean|sd)$"
  ) %>%
  pivot_wider(
    names_from = year,
    values_from = value
  ) %>% 
  mutate(across(where(is.numeric), ~round(.x, 2)))

```


## Control variables 

###  Percent of adults 25 and over with a Bachelor’s degree or higher by state in 2018 (numeric: %, 0 to 100) -- ACS

```{r}
acs_2018 <- import(here("Data/Non-IPEDS Data/ACS Educational Attainment by State/ACSST1Y2018.S1501-2025-04-09T015452.csv"))

stabbr_csv <- import(here("Data/Non-IPEDS Data/states.csv"))

educ_attain_2018 <- select(acs_2018, `Label (Grouping)`, ends_with("Percent!!Estimate"))[16,] %>%
  pivot_longer(cols = 2:53, names_to = "ACS_variable") %>%
  mutate(
    State = str_extract(string = ACS_variable, pattern = "^[^!]+(?=!)")
  ) %>%
  left_join(
    y = stabbr_csv,
    by = "State"
  ) %>%
  rename(ACS_Pct_Bachelors = value) %>%
  mutate(ACS_Pct_Bachelors = as.numeric(str_extract(ACS_Pct_Bachelors, pattern="\\d+\\.\\d+")))

educ_attain_clean <- educ_attain_2018 %>% 
  janitor::clean_names() %>% 
  select(acs_pct_bachelors, abbreviation) %>% 
  rename(stabbr = abbreviation,
         pct_bachelors_2018 = acs_pct_bachelors)
```


###  High-school graduates by state (numeric: integer) -- WICHE

```{r}
wiche_data_allrows <- import(here("Data/Non-IPEDS Data/WICHE HS Graduates by State.xlsx"), which = "Data")


hs_grads_by_year <- wiche_data_allrows %>%
  filter(
    Grade == "High school graduates", # Keep only rows ~ total high school graduates
    RaceEthnicity == "Total/any", # Keep only rows ~ total all race/ethnicity
    SchoolSector == "Grand Total (public+private)", # Keep only rows ~ total public and private high schools
    !(str_detect(string = Stabbr, pattern = "^_")), # Keep only rows NOT starting with an underscore (i.e., drop region aggregates)
    !(str_detect(string = Stabbr, pattern = "\\d")), # Keep only rows NOT containing digits (i.e., drop region aggregates)
    !(Stabbr %in% c("CM","CW")) # Keep only rows NOT CM and CW (i.e., drop region aggregates)
  ) %>%
  mutate(
    HS_Grad_Year = as.numeric(str_extract(string = ClassOf, pattern = "\\d+")) # Extract the graduation year for numeric filtering
  ) %>%
  filter(
    HS_Grad_Year >= 2017,
    HS_Grad_Year < 2026
  ) %>%
  select(
    # Keep only these variables
    Stabbr,
    HS_Grad_Year,
    Students
  ) %>%
  # Convert to wide format...
  pivot_wider(id_cols = Stabbr, names_from = HS_Grad_Year, values_from = Students, names_prefix = "HS_Grad_") %>% 
  janitor::clean_names() %>% 
  select(stabbr, hs_grad_2018:hs_grad_2022)

# make join-friendly
hs_grads_clean <- hs_grads_by_year %>% 
  pivot_longer(
    cols = starts_with("hs_grad_"),
    names_to = "year",
    names_prefix = "hs_grad_",
    values_to = "hs_grads"
  ) %>%
  mutate(year = as.integer(year) + 1) #this is actually the HS grad for the prev year, but needed to add 1 to join with full data
```

```{r}
state_dat <- left_join(hs_grads_clean, educ_attain_clean)
```


###  Degree of urbanization (categorical: city/suburb, town/rural) 

```{r}
ipeds_geog_values <- import(
  here("Data/Data Geographic Variables/Data Geographic Variables.csv")
)[, 1:9]

ipeds_geog_labels <- import(here("Data/Data Geographic Variables/Labels Geographic Variables.csv"))

ipeds_geog_values$Locale_Rural <- case_when(
  ipeds_geog_values$`Degree of urbanization (Urban-centric locale) (HD2023)` %in% c(11,12,13,21,22,23) ~ "City/Suburb",
  ipeds_geog_values$`Degree of urbanization (Urban-centric locale) (HD2023)` %in% c(31,32,33,41,42,43) ~ "Town/Rural",
  TRUE ~ NA
)
```

```{r}
geo <- ipeds_geog_values %>% 
  janitor::clean_names() %>% 
  select(unit_id, institution_name, state_abbreviation_hd2023, locale_rural) %>% 
  rename(stabbr = state_abbreviation_hd2023) %>% 
  left_join(., state_dat) %>% 
  rename(hs_grad_state = hs_grads,
         pct_bachelors_2018_state = pct_bachelors_2018)

head(geo)
```

```{r}
describe(geo)
```

Joined dataset
```{r}
full_dat <- left_join(dat, geo)
```

```{r}
describe(full_dat)
```


###  Carnegie ranking (categorical: R1, R2, M, other) 



