mo---
title: "Final Analysis & Interpretation"
output: html_notebook
---

# Setup

```{r}
library(tidyverse)
library(lme4)
```

```{r}
full_dat3 <- read_csv("../Data/full_dat3.csv")
# imputed_enrl <- read_csv("../Data/Data Sector-Enrollment Variables/Imputation_Flags.csv")
# enrl_data <- read_csv("../Data/final_enrollment_data.csv")
```

```{r}
# enrl_long <- enrl_data %>%
#   select(-`...18`) %>%
#   pivot_longer(cols = 3:17, names_to = "Var_Name", values_to = "IPEDS_Value") %>%
#   mutate(
#     Year = str_extract(Var_Name, "20\\d\\d"),
#     Enrl_Var = case_when(
#       str_detect(Var_Name, "^Undergraduate") ~ "UGRD_ENRL",
#       str_detect(Var_Name, "^First") ~ "FYR_ENRL",
#       str_detect(Var_Name, "Total") ~ "TOT_ENRL"
#     )
#   )
# 
# enrl_long %>%
#   group_by(UnitID, Enrl_Var) %>%
#   summarise(N = sum(!is.na(IPEDS_Value))) %>%
#   ungroup() %>%
#   pivot_wider(names_from = "Enrl_Var", values_from = "N") %>%
#   filter(FYR_ENRL < 5 | TOT_ENRL < 5 | UGRD_ENRL < 5) %>%
#   left_join(
#     y = imputed_enrl,
#     by = "UnitID"
#   )

```


```{r}
combined_data <- full_dat3 %>%
  # Inclusion criteria:
  # In the Carnegie universe
  filter(carnegie != "Not Carnegie") %>%
  # At least 1000 overall enrollment in first reporting year
  group_by(unit_id) %>%
  filter(year == min(year, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(overall_enrollment >= 1000) %>%
  distinct(unit_id) %>%
  left_join(
    y = full_dat3,
    by = "unit_id"
  ) %>%
  # How many years of data?
  group_by(unit_id) %>%
  summarise(N_year = length(unique(year))) %>%
  ungroup() %>%
  left_join(
    y = full_dat3,
    by = "unit_id"
  )

public_dat3 <- combined_data %>%
  filter(sector=="public") %>%
  mutate(
    year_0 = year - 2019,
    # Log transforms: first-time enrollment; HS grads by state; applicants; average net price with aid
    log_fyr = log(first_time_enrollment),
    log_hs = log(hs_grad_state),
    log_apps = log(applicants),
    log_price = log(avg_net_price_w_aid),
    sy_fyr = scale(log_fyr),
    x_year = year_0,
    sxl_hs = scale(log_hs), # log this
    sx_bach = scale(pct_bachelors_2018_state),
    x_rural = factor(locale_rural),
    sx_pell = scale(pct_pell),
    sx_race_minority = scale(pct_race_minority),
    sx_price = scale(avg_net_price_w_aid), # don't log this
    sxl_apps = scale(log_apps), # log this
    sx_admit_rate = scale(admissions_rate)
  ) %>%
  mutate(
    carnegie_category = factor(
      carnegie,
      levels = c("R1", "R2", "R0", "M", "Baccalaureate", "Special or Tribal"),
      labels = c("R1", "R2", "D/PU", "M", "B", "SF&TCU")
    )
  ) %>%
  left_join(
    y = combined_data %>%
      group_by(unit_id) %>%
      filter(year == min(year)) %>%
      ungroup() %>%
      mutate(
        size_category = case_when(
          overall_enrollment >= 1000 & overall_enrollment < 5000 ~ "1k-5k",
          overall_enrollment >= 5000 & overall_enrollment < 10000 ~ "5k-10k",
          overall_enrollment >= 10000 & overall_enrollment < 20000 ~ "10k-20k",
          overall_enrollment >= 20000 ~ "20k+"
        ) %>% factor(levels = c("1k-5k","5k-10k","10k-20k","20k+"))
      ) %>%
      select(unit_id, size_category),
    by = "unit_id"
  )

private_dat3 <- combined_data %>%
  filter(sector=="private") %>%
  mutate(
    year_0 = year - 2019,
    # Log transforms: first-time enrollment; HS grads by state; applicants; average net price with aid
    log_fyr = log(first_time_enrollment),
    log_hs = log(hs_grad_state),
    log_apps = log(applicants),
    log_price = log(avg_net_price_w_aid),
    sy_fyr = scale(log_fyr),
    x_year = year_0,
    sxl_hs = scale(log_hs), # log this
    sx_bach = scale(pct_bachelors_2018_state),
    x_rural = factor(locale_rural),
    sx_pell = scale(pct_pell),
    sx_race_minority = scale(pct_race_minority),
    sx_price = scale(avg_net_price_w_aid), # don't log this
    sxl_apps = scale(log_apps), # log this
    sx_admit_rate = scale(admissions_rate)
  ) %>%
  mutate(
    carnegie_category = factor(
      carnegie,
      levels = c("R1", "R2", "R0", "M", "Baccalaureate", "Special or Tribal"),
      labels = c("R1", "R2", "D/PU", "M", "B", "SF&TCU")
    )
  ) %>%
  left_join(
    y = combined_data %>%
      group_by(unit_id) %>%
      filter(year == min(year)) %>%
      ungroup() %>%
      mutate(
        size_category = case_when(
          overall_enrollment >= 1000 & overall_enrollment < 5000 ~ "1k-5k",
          overall_enrollment >= 5000 & overall_enrollment < 10000 ~ "5k-10k",
          overall_enrollment >= 10000 & overall_enrollment < 20000 ~ "10k-20k",
          overall_enrollment >= 20000 ~ "20k+"
        ) %>% factor(levels = c("1k-5k","5k-10k","10k-20k","20k+"))
      ) %>%
      select(unit_id, size_category),
    by = "unit_id"
  )


# full_dat3 %>%
#   select(unit_id, year, overall_enrollment, first_time_enrollment) %>%
#   pivot_wider(names_from = "year", id_cols = "unit_id", values_from = c("overall_enrollment","first_time_enrollment"))
# 
# instid <- unique(full_dat3$unit_id)
# writeClipboard(paste0(instid[1:500], collapse=","))
# writeClipboard(paste0(instid[501:1000], collapse=","))
# writeClipboard(paste0(instid[1001:1500], collapse = ","))
# writeClipboard(paste0(instid[1501:length(instid)], collapse=","))
```


```{r}
formula_1 <- formula(
  first_time_enrollment ~
    1
    + x_year
    # + (1 + x_year | carnegie)
    + (1 + x_year | unit_id)
)

formula_2 <- formula(
  first_time_enrollment ~
    1
    + x_year
    + x_rural
    + sx_bach # Percent of adults with bachelors degree in the state where the institution is located, standardized
    + sxl_hs # The number of high school graduates that year, logged and standardized
    # + (1 + x_year | carnegie)
    + (1 + x_year | unit_id)
)

formula_3 <- formula(
  first_time_enrollment ~
    1
    + x_year
    + x_rural
    + sx_bach # Percent of adults with bachelors degree in the state where the institution is located, standardized
    + sxl_hs # The number of high school graduates that year, logged and standardized
    + sx_race_minority
    + sx_pell
    + sxl_apps
    + sx_admit_rate
    + sx_price
    # + (1 + x_year | carnegie)
    + (1 + x_year | unit_id)
)
```


```{r}
glm_public_1 <- glmer(formula = formula_1, data = public_dat3, family = "poisson")
glm_public_2 <- glmer(formula = formula_2, data = public_dat3, family = "poisson")
glm_public_3 <- glmer(formula = formula_3, data = public_dat3, family = "poisson")

glm_private_1 <- glmer(formula = formula_1, data = private_dat3, family = "poisson")
glm_private_2 <- glmer(formula = formula_2, data = private_dat3, family = "poisson")
glm_private_3 <- glmer(formula = formula_3, data = private_dat3, family = "poisson")

```

```{r}
map(
  .x = list(
    glm_public_1,
    glm_public_2,
    glm_public_3,
    glm_private_1,
    glm_private_2,
    glm_private_3
  ),
  .f = summary
)
```

Creating tables -- public
```{r}
public_mods <- list(glm_public_1, glm_public_2, glm_public_3)

## table creation
modelsummary::modelsummary(public_mods, 
                           title = "Model Summary for Public Institutions",
                           notes = "***p<.01, **p<.05, *p<.10",
                           estimate = "{estimate}{stars}({std.error})",
                           statistic = NULL,
                           coef_rename = c(
  "x_year" = "Year", 
  "x_ruralTown/Rural" = "Town/Rural",
  "sx_bach" = "Pct. Bachelor-degree Holders",
  "sxl_hs" = "HS Graduates",
  "sx_race_minority" = "Pct. Non-White",
  "sx_pell" = "Pct. Pell",
  "sxl_apps" = "Applicants",
  "sx_admit_rate" = "Admission Rate",
  "sx_price" = "Tuition Cost"),
  coef_omit = "SD|Cor|Intercept.*unit_id|x_year.*unit_id",
  gof_map = c("nobs", "aic", "bic", "rmse", "logLik"))
```

Private mods
```{r}
private_mods <- list(glm_private_1, glm_private_2, glm_private_3)

## table creation
modelsummary::modelsummary(private_mods, 
                           output = "gt",
                           title = "***p<.01, **p<.05, *p<.10",
                           estimate = "{estimate}{stars}({std.error})",
                           statistic = NULL,
                           coef_rename = c(
  "x_year" = "Year", 
  "x_ruralTown/Rural" = "Town/Rural",
  "sx_bach" = "Pct. Bachelor-degree Holders",
  "sxl_hs" = "HS Graduates",
  "sx_race_minority" = "Pct. Non-White",
  "sx_pell" = "Pct. Pell",
  "sxl_apps" = "Applicants",
  "sx_admit_rate" = "Admission Rate",
  "sx_price" = "Tuition Cost"),
  coef_omit = "SD|Cor|Intercept.*unit_id|x_year.*unit_id",
  gof_map = c("nobs", "aic", "bic", "rmse", "logLik")) %>% 
  gt::tab_spanner(label = "Private", columns = everything())
```

All in one table
```{r}
all_mods <- list(glm_public_1, glm_public_2, glm_public_3, glm_private_1, glm_private_2, glm_private_3)

summary_table <- modelsummary::modelsummary(all_mods, 
                           output = "gt",
                           title = "Model Summaries on Enrollment Rate by Institution Type",
                           notes = "***p<.01, **p<.05, *p<.10",
                           estimate = "{estimate}{stars}({std.error})",
                           statistic = NULL,
                           coef_rename = c(
  "x_year" = "Year", 
  "x_ruralTown/Rural" = "Town/Rural (compared to City/Suburb)",
  "sx_bach" = "Pct. Bachelor-degree Holders",
  "sxl_hs" = "HS Graduates",
  "sx_race_minority" = "Pct. Non-White",
  "sx_pell" = "Pct. Pell",
  "sxl_apps" = "Applicants",
  "sx_admit_rate" = "Admission Rate",
  "sx_price" = "Avg. Net Price with Aid"),
  coef_omit = "SD|Cor|Intercept.*unit_id|x_year.*unit_id",
  gof_map = c("nobs", "aic", "bic", "rmse", "logLik")) %>% 
  gt::tab_spanner(label = "Public", columns = 2:4) %>% 
  gt::tab_spanner(label = "Private", columns = 5:7)
```


```{r}
anova(glm_public_1, glm_public_2, glm_public_3)
anova(glm_private_1, glm_private_2, glm_private_3)
```

```{r}
public_intslopes <- tibble(
  unit_id = rownames(coef(glm_public_1)$unit_id) %>% as.numeric(),
  m1_int = coef(glm_public_1)$unit_id[,"(Intercept)"],
  m1_slope_time = coef(glm_public_1)$unit_id[,"x_year"]
) %>%
  left_join(
    y = tibble(
      unit_id = rownames(coef(glm_public_2)$unit_id) %>% as.numeric(),
      m2_int = coef(glm_public_2)$unit_id[,"(Intercept)"],
      m2_slope_time = coef(glm_public_2)$unit_id[,"x_year"]
    ),
    by = "unit_id"
  ) %>%
  left_join(
    y = tibble(
      unit_id = rownames(coef(glm_public_3)$unit_id) %>% as.numeric(),
      m3_int = coef(glm_public_3)$unit_id[,"(Intercept)"],
      m3_slope_time = coef(glm_public_3)$unit_id[,"x_year"]
    ),
    by = "unit_id"
  )

# public_dat3$m1_int <- coef(glm_public_1)$unit_id[,"(Intercept)"]
# public_dat3$m1_slope_time <- coef(glm_public_1)$unit_id[,"x_year"]
public_dat3$m1_fitted_link <- predict(glm_public_1, type="link")
public_dat3$m1_fitted_resp <- predict(glm_public_1, type="resp")
public_dat3$m1_residual <- resid(glm_public_1)
# public_dat3$m2_int <- coef(glm_public_2)$unit_id[,"(Intercept)"]
# public_dat3$m2_slope_time <- coef(glm_public_2)$unit_id[,"x_year"]
public_dat3$m2_fitted_link <- predict(glm_public_2, type="link")
public_dat3$m2_fitted_resp <- predict(glm_public_2, type="resp")
public_dat3$m2_residual <- resid(glm_public_2)
# public_dat3$m3_int <- coef(glm_public_3)$unit_id[,"(Intercept)"]
# public_dat3$m3_slope_time <- coef(glm_public_3)$unit_id[,"x_year"]
public_dat3$m3_fitted_link <- predict(glm_public_3, type="link")
public_dat3$m3_fitted_resp <- predict(glm_public_3, type="resp")
public_dat3$m3_residual <- resid(glm_public_3)

#

private_intslopes <- tibble(
  unit_id = rownames(coef(glm_private_1)$unit_id) %>% as.numeric(),
  m1_int = coef(glm_private_1)$unit_id[,"(Intercept)"],
  m1_slope_time = coef(glm_private_1)$unit_id[,"x_year"]
) %>%
  left_join(
    y = tibble(
      unit_id = rownames(coef(glm_private_2)$unit_id) %>% as.numeric(),
      m2_int = coef(glm_private_2)$unit_id[,"(Intercept)"],
      m2_slope_time = coef(glm_private_2)$unit_id[,"x_year"]
    ),
    by = "unit_id"
  ) %>%
  left_join(
    y = tibble(
      unit_id = rownames(coef(glm_private_3)$unit_id) %>% as.numeric(),
      m3_int = coef(glm_private_3)$unit_id[,"(Intercept)"],
      m3_slope_time = coef(glm_private_3)$unit_id[,"x_year"]
    ),
    by = "unit_id"
  )

# private_dat3$m1_int <- coef(glm_private_1)$unit_id[,"(Intercept)"]
# private_dat3$m1_slope_time <- coef(glm_private_1)$unit_id[,"x_year"]
private_dat3$m1_fitted_link <- predict(glm_private_1, type="link")
private_dat3$m1_fitted_resp <- predict(glm_private_1, type="resp")
private_dat3$m1_residual <- resid(glm_private_1)
# private_dat3$m2_int <- coef(glm_private_2)$unit_id[,"(Intercept)"]
# private_dat3$m2_slope_time <- coef(glm_private_2)$unit_id[,"x_year"]
private_dat3$m2_fitted_link <- predict(glm_private_2, type="link")
private_dat3$m2_fitted_resp <- predict(glm_private_2, type="resp")
private_dat3$m2_residual <- resid(glm_private_2)
# private_dat3$m3_int <- coef(glm_private_3)$unit_id[,"(Intercept)"]
# private_dat3$m3_slope_time <- coef(glm_private_3)$unit_id[,"x_year"]
private_dat3$m3_fitted_link <- predict(glm_private_3, type="link")
private_dat3$m3_fitted_resp <- predict(glm_private_3, type="resp")
private_dat3$m3_residual <- resid(glm_private_3)

```

Growth and decline summary table
```{r}
public_intslopes <- public_intslopes %>%
  mutate(
    Change_Sign = case_when(
      m1_slope_time < 0 ~ "Negative"
      ,m1_slope_time >= 0 ~ "Positive"
    )
    ,Change_Category = case_when(
      m1_slope_time <= -0.025 ~ "Decline (>2.5%)"
      ,m1_slope_time > -0.025 & m1_slope_time < 0 ~ "Stable (Negative, <2.5%)"
      ,m1_slope_time >= 0 & m1_slope_time < 0.025 ~ "Stable (Positive, <2.5%)"
      ,m1_slope_time >= 0.025 ~ "Growth (>2.5%)"
    )
  ) %>%
  left_join(
    y = public_dat3 %>% distinct(unit_id, size_category, carnegie_category),
    by = "unit_id"
  )

private_intslopes <- private_intslopes %>%
  mutate(
    Change_Sign = case_when(
      m1_slope_time < 0 ~ "Negative"
      ,m1_slope_time >= 0 ~ "Positive"
    )
    ,Change_Category = case_when(
      m1_slope_time <= -0.025 ~ "Decline (>2.5%)"
      ,m1_slope_time > -0.025 & m1_slope_time < 0 ~ "Stable (Negative, <2.5%)"
      ,m1_slope_time >= 0 & m1_slope_time < 0.025 ~ "Stable (Positive, <2.5%)"
      ,m1_slope_time >= 0.025 ~ "Growth (>2.5%)"
    )
  ) %>%
  left_join(
    y = private_dat3 %>% distinct(unit_id, size_category, carnegie_category),
    by = "unit_id"
  )
```

```{r}
combined_change <- rbind(
  public_intslopes %>% mutate(sector = "public"),
  private_intslopes %>% mutate(sector = "private")
)

xtabs_change <- table(
  combined_change$Change_Category,
  combined_change$sector
)[c("Decline (>2.5%)","Stable (Negative, <2.5%)","Stable (Positive, <2.5%)","Growth (>2.5%)"),c("public","private")]

xtabs_change
marginSums(xtabs_change, 1)
marginSums(xtabs_change, 2)
xtabs_change[,1]/marginSums(xtabs_change, 2)[1]
xtabs_change[,2]/marginSums(xtabs_change, 2)[2]
marginSums(xtabs_change, 1)/sum(xtabs_change)

#table(combined_change)
```


```{r}
# Outliers
public_outliers <- public_intslopes %>%
  mutate(
    m1_outlier_slope = case_when(
      m1_slope_time < -3 * sd(m1_slope_time) + mean(m1_slope_time) ~ "Outlier Decline",
      m1_slope_time > 3 * sd(m1_slope_time) + mean(m1_slope_time) ~ "Outlier Growth"
    )
  ) %>%
  filter(!is.na(m1_outlier_slope)) %>%
  left_join(
    y = public_dat3,
    by = "unit_id"
  )

private_outliers <- private_intslopes %>%
  mutate(
    m1_outlier_slope = case_when(
      m1_slope_time < -3 * sd(m1_slope_time) + mean(m1_slope_time) ~ "Outlier Decline",
      m1_slope_time > 3 * sd(m1_slope_time) + mean(m1_slope_time) ~ "Outlier Growth"
    )
  ) %>%
  filter(!is.na(m1_outlier_slope)) %>%
  left_join(
    y = private_dat3,
    by = "unit_id"
  )

write_csv(public_outliers %>% map_df(.f = function(x) if(is.matrix(x)) {x[,1]} else {x}), "../Results/Public_Outliers.csv")
write_csv(private_outliers %>% map_df(.f = function(x) if(is.matrix(x)) {x[,1]} else {x}), "../Results/Private_Outliers.csv")

```

# Visualization

```{r}
ggplot(
  data = public_dat3,
  mapping = aes(x = x_year, y = m1_fitted_resp)
) +
  geom_line(mapping = aes(group=unit_id))
```

## Slope change from model 1 to model 2

```{r}
public_m1m2_slopechange <- public_intslopes %>%
  mutate(slopechange_21 = m2_slope_time - m1_slope_time) %>%
  arrange(desc(slopechange_21)) %>%
  mutate(unit_id_f = factor(unit_id, levels = unique(unit_id))) %>%
  left_join(
    y = distinct(public_dat3, unit_id, carnegie_category, size_category),
    by = "unit_id"
  )

private_m1m2_slopechange <- private_intslopes %>%
  mutate(slopechange_21 = m2_slope_time - m1_slope_time) %>%
  arrange(desc(slopechange_21)) %>%
  mutate(unit_id_f = factor(unit_id, levels = unique(unit_id))) %>%
  left_join(
    y = distinct(private_dat3, unit_id, carnegie_category, size_category),
    by = "unit_id"
  )


ggplot(data=public_m1m2_slopechange, mapping = aes(x=unit_id_f, y=slopechange_21)) +
  geom_bar(
    stat="identity"
#    ,color = "black"
    ,mapping = aes(fill=size_category)
  ) +
  scale_fill_brewer(type = "div", direction = -1) +
#  facet_wrap(~size_category) +
  theme_classic() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

public_m1m2_slopechange %>%
  mutate(
    slpchg_median = median(slopechange_21),
    slpchg_top3rd = quantile(slopechange_21, 0.66),
    slpchg_bot3rd = quantile(slopechange_21, 0.33)
  ) %>%
  group_by(size_category, carnegie_category) %>%
  summarise(
    Top3rd = sum(slopechange_21 > slpchg_top3rd),
    Bot3rd = -1*sum(slopechange_21 < slpchg_bot3rd)
  ) %>%
  pivot_longer(cols=3:4) %>%
  ggplot(aes(x = size_category, y = value)) +
  geom_bar(stat="identity", mapping=aes(fill=name)) +
  facet_wrap(~carnegie_category)

ggplot(
  public_m1m2_slopechange,
  aes(x = m1_slope_time, y = m2_slope_time)
) +
  geom_point()

set.seed(3)
slp21_example <- sample_n(public_m1m2_slopechange, 25)

public_dat3 %>%
  filter(unit_id %in% slp21_example$unit_id) %>%
  ggplot(
    mapping = aes(x = year_0)
  ) +
  geom_line(
    mapping = aes(y = m2_fitted_resp, group = unit_id)
    ,color="red",linewidth=2,alpha=0.2
  ) +
  geom_line(
    mapping = aes(y = m1_fitted_resp, group = unit_id)
    ,color="blue",linewidth=2,alpha=0.2
  ) +
  facet_wrap(~unit_id, nrow=5)

public_dat3 %>%
  filter(unit_id %in% slp21_example$unit_id) %>%
  ggplot(
    mapping = aes(x = year_0)
  ) +
  geom_line(
    mapping = aes(y = m2_fitted_link, group = unit_id)
    ,color="red",linewidth=2,alpha=0.2
  ) +
  geom_line(
    mapping = aes(y = m1_fitted_link, group = unit_id)
    ,color="blue",linewidth=2,alpha=0.2
  ) +
  facet_wrap(~unit_id, nrow=5)
```




# Within-Institution Correlation...

```{r}
tibble(
  unit_id = unique(public_dat3$unit_id)
  ,enrl_cor_hs = map_dbl(
    .x = as.list(unique(public_dat3$unit_id)),
    .f = function(x) {
      unit_id_subset <- filter(public_dat3, unit_id==x) %>%
        arrange(year)
    
      cor(unit_id_subset$first_time_enrollment, unit_id_subset$hs_grad_state)
    }
  )
  ,enrl_cor_pell = map_dbl(
    .x = as.list(unique(public_dat3$unit_id)),
    .f = function(x) {
      unit_id_subset <- filter(public_dat3, unit_id==x) %>%
        arrange(year)
    
      cor(unit_id_subset$first_time_enrollment, unit_id_subset$pct_pell)
    }
  )
) %>%
  pivot_longer(cols = 2:ncol(.)) %>%
  ggplot(aes(x=name, y=value)) +
  geom_boxplot()

```

```{r}
m1_dist <- dist(public_intslopes %>% select(m1_int, m1_slope_time))
```

```{r}
set.seed(20250514)
m1_trj_medoids <- kmed::rankkmed(m1_dist, ncluster = 7)
public_intslopes[m1_trj_medoids$medoid,] %>%
  left_join(y=public_dat3,
            by="unit_id") %>%
  ggplot(aes(x=year_0, y=log_fyr)) +
  geom_line(mapping=aes(group=unit_id,color=factor(unit_id)), linewidth=2, linetype=1, alpha=0.3) +
  geom_abline(aes(slope=m1_slope_time, intercept = m1_int, color=factor(unit_id)), linewidth=2) +
  theme_classic()


m1_trj_medoids$cluster
# public_intslopes %>%
#   mutate(
#     trj_cluster=m1_trj_medoids$cluster,
#     rowind = row_number(),
#     is_medoid = if_else(rowind %in% m1_trj_medoids$medoid, TRUE, FALSE)
#   ) %>%
#   left_join(y=public_dat3, by="unit_id") %>%
#   ggplot(
#     data = .,
#     aes(x=year_0,y=log_fyr)
#   ) +
#   geom_line(mapping=aes(group=unit_id), linewidth=2, linetype=1, alpha=0.3) +
#   geom_abline(mapping = aes(slope=m1_slope_time, intercept = m1_int, color=factor(trj_cluster), group=unit_id))
```


```{r}
public_fyr_change_5yr <- public_dat3 %>%
  select(
    unit_id, sector, year_0, first_time_enrollment, log_fyr, m1_fitted_link, m1_fitted_resp
  ) %>%
  group_by(unit_id, sector) %>%
  filter(year_0 == max(year_0) | year_0 == min(year_0)) %>%
  mutate(year_ref = case_when(year_0==max(year_0)~"LastYr", year_0==min(year_0)~"FirstYr")) %>%
  pivot_wider(id_cols = c(unit_id, sector), names_from = year_ref, values_from=c(m1_fitted_resp,first_time_enrollment)) %>%
  mutate(
    MaxMinYr_Change_Fit = (m1_fitted_resp_LastYr/m1_fitted_resp_FirstYr)-1
    ,MaxMinYr_Change_Obs = (first_time_enrollment_LastYr/first_time_enrollment_FirstYr)-1
  ) %>%
  mutate(
    Growth_Category = case_when(
      MaxMinYr_Change_Fit < -0.05 ~ "Decline more than 5%",
      MaxMinYr_Change_Fit >= -0.05 & MaxMinYr_Change_Fit <= 0.05 ~ "Change less than 5%",
      MaxMinYr_Change_Fit > 0.05 ~ "Growth more than 5%"
    )
  ) %>%
  ungroup()

private_fyr_change_5yr <- private_dat3 %>%
 select(
    unit_id, sector, year_0, first_time_enrollment, log_fyr, m1_fitted_link, m1_fitted_resp
  ) %>%
  group_by(unit_id, sector) %>%
  filter(year_0 == max(year_0) | year_0 == min(year_0)) %>%
  mutate(year_ref = case_when(year_0==max(year_0)~"LastYr", year_0==min(year_0)~"FirstYr")) %>%
  pivot_wider(id_cols = c(unit_id, sector), names_from = year_ref, values_from=c(m1_fitted_resp,first_time_enrollment)) %>%
  mutate(
    MaxMinYr_Change_Fit = (m1_fitted_resp_LastYr/m1_fitted_resp_FirstYr)-1
    ,MaxMinYr_Change_Obs = (first_time_enrollment_LastYr/first_time_enrollment_FirstYr)-1
  ) %>%
  mutate(
    Growth_Category = case_when(
      MaxMinYr_Change_Fit < -0.05 ~ "Decline more than 5%",
      MaxMinYr_Change_Fit >= -0.05 & MaxMinYr_Change_Fit <= 0.05 ~ "Change less than 5%",
      MaxMinYr_Change_Fit > 0.05 ~ "Growth more than 5%"
    )
  ) %>%
  ungroup()
```

```{r}
rbind(
  public_fyr_change_5yr,
  private_fyr_change_5yr
) %>%
  count(sector, Growth_Category) %>%
  group_by(sector) %>%
  mutate(pct = round(100*n/sum(n))) %>%
  arrange(sector, desc(n))
```
```{r}
public_intslopes[m1_trj_medoids$medoid,] %>%
  left_join(y=public_dat3,
            by="unit_id") %>%
  ggplot(aes(x=year_0, y=log_fyr)) +
  geom_line(mapping=aes(group=unit_id,color=factor(unit_id)), linewidth=2, linetype=1, alpha=0.3) +
  # geom_abline(aes(slope=m1_slope_time, intercept = m1_int, color=factor(unit_id)), linewidth=2) +
  geom_abline(aes(slope=m2_slope_time, intercept = m2_int, color=factor(unit_id)), linewidth=2, linetype=2) +
  scale_y_continuous(limits = c(5.5, 8)) +
  theme_classic()
```
```{r}
public_dat3 %>%
  distinct(stabbr, year, hs_grad_state, pct_bachelors_2018_state, sxl_hs, sx_bach) %>%
  arrange(sx_bach, stabbr, year)

summary(glm_public_2)
```


MS, TX, NE, CA, VT
```{r}
public_dat3 %>%
  distinct(stabbr, year, hs_grad_state, pct_bachelors_2018_state, sxl_hs, sx_bach) %>%
  arrange(sx_bach, stabbr, year) %>%
  filter(stabbr %in% c("MS","TX","NE","CA","VT")) %>%
  mutate(
    State_Link = glm_public_2@beta[1] + glm_public_2@beta[2]*(year-2019) + glm_public_2@beta[4]*sx_bach + glm_public_2@beta[5]*sxl_hs,
    State_Resp = exp(State_Link)
  )
summary(glm_public_2)
```

```{r}
public_dat3 %>%
  distinct(stabbr, year, hs_grad_state, pct_bachelors_2018_state, sxl_hs, sx_bach) %>%
  arrange(sx_bach, stabbr, year) %>%
  ggplot(aes(x=pct_bachelors_2018_state, y=hs_grad_state, color=stabbr)) +
  geom_point()

state_summary <- public_dat3 %>%
  distinct(stabbr, year, hs_grad_state, pct_bachelors_2018_state) %>%
  group_by(stabbr) %>%
  summarise(
    Avg_HS_Grads = mean(hs_grad_state),
    Pct_Bach = mean(pct_bachelors_2018_state)
  ) %>%
  ungroup() %>%
  mutate(
    SL_HS_Grads = scale(log(Avg_HS_Grads)),
    HS_Grads_Rank = rank(Avg_HS_Grads),
    Pct_Bach_Rank = rank(Pct_Bach)
  ) %>%
  arrange(desc(Pct_Bach)) %>%
  mutate(State_Ordered = factor(stabbr, levels=unique(stabbr)))

ggplot(data=state_summary, mapping=aes(x=State_Ordered, y=Pct_Bach)) +
  geom_bar(stat="identity", mapping=aes(fill=SL_HS_Grads)) +
  scale_fill_distiller(type="div") +
  theme_classic()
```
Putting things back to real-world numbers

```{r}
# Towns and Rural Areas (Model 2), Public and Private
# Public: B = -0.711 per unit increase in x_rural (0 = Urban, 1 = Rural)
# Private: B = -0.258 per unit increase in x_rural (0 = Urban, 1 = Rural)
exp(-0.711) ; exp(-0.258)

# High-school graduates (Model 2), Public and Private
# Public: B = 0.212 per unit increase in sxl_hs (0 = 11.35922, 1 = 12.35907, s = 0.9998539)
public_dat3$sxl_hs %>% attr("scaled:center") ; public_dat3$sxl_hs %>% attr("scaled:scale")
# Double the high-school graduates
exp(0.212 * (log(2*exp(public_dat3$sxl_hs %>% attr("scaled:center"))) - (public_dat3$sxl_hs %>% attr("scaled:center")))/(public_dat3$sxl_hs %>% attr("scaled:scale")))
# Private: B = 0.081 per unit increase in sxl_hs (0 = 11.49692, 1 = 12.38718, s = )
private_dat3$sxl_hs %>% attr("scaled:center") ; private_dat3$sxl_hs %>% attr("scaled:scale")
# Double the high-school graduates
exp(0.081 * (log(2*exp(private_dat3$sxl_hs %>% attr("scaled:center"))) - (private_dat3$sxl_hs %>% attr("scaled:center")))/(private_dat3$sxl_hs %>% attr("scaled:scale")))

# Educational attainment (Model 2), Public and Private
# Public: B = -0.112 per unit increase in sx_bach (0 = 32.18005, 1 = 37.52799, s = 5.347944)
attr(public_dat3$sx_bach, "scaled:center"); attr(public_dat3$sx_bach, "scaled:scale")
# Increase high-school graduates by one standard deviation
1/exp(-0.112)
# Private: B = 0.138 per unit increase in sx_bach (0 = 33.50603, 1 = 39.17788, s = 5.671852)
attr(private_dat3$sx_bach, "scaled:center"); attr(private_dat3$sx_bach, "scaled:scale")
exp(0.138)

# Application volume (Model 3), Public and Private
# Public: B = 0.514 per unit increase in sxl_apps (0 = 8.954673, 1 = 10.13351, s = 1.17884)
attr(public_dat3$sxl_apps, "scaled:center"); attr(public_dat3$sxl_apps, "scaled:scale")
# Increase apps by 10 percent
exp(0.514*(log(1.1*exp(attr(public_dat3$sxl_apps, "scaled:center"))) - attr(public_dat3$sxl_apps, "scaled:center"))/(attr(public_dat3$sxl_apps, "scaled:scale")))
# Private: B = 0.469 per unit increase in sxl_apps (0 = 8.306277, 1 = 9.345809, s = 1.039532)
attr(private_dat3$sxl_apps, "scaled:center") ; attr(private_dat3$sxl_apps, "scaled:scale")
# Increase apps by 10 percent
exp(0.469*(log(1.1*exp(attr(private_dat3$sxl_apps, "scaled:center"))) - attr(private_dat3$sxl_apps, "scaled:center"))/(attr(private_dat3$sxl_apps, "scaled:scale")))

# Admit rate (Model 3), Public and Private
# Public: B = 0.114 per unit increase in sx_admit_rate (0 = 0.765045, 1 = 0.9392904, s = 0.1742454)
attr(public_dat3$sx_admit_rate, "scaled:center"); attr(public_dat3$sx_admit_rate, "scaled:scale")
# Increase from 80% (median) to 85% admission rate
exp(0.114 * ((0.85 - attr(public_dat3$sx_admit_rate, "scaled:center"))/(attr(public_dat3$sx_admit_rate, "scaled:scale")) - (0.80 - attr(public_dat3$sx_admit_rate, "scaled:center"))/(attr(public_dat3$sx_admit_rate, "scaled:scale"))))
# Private: B = 0.162 per unit increase in sx_admit_rate (0 = 0.6748105, 1 = 0.9032039, s=0.2283934)
attr(private_dat3$sx_admit_rate, "scaled:center"); attr(private_dat3$sx_admit_rate, "scaled:scale")
# Increase from 80% (median) to 85% admission rate
exp(0.162 * ((0.85 - attr(private_dat3$sx_admit_rate, "scaled:center"))/(attr(private_dat3$sx_admit_rate, "scaled:scale")) - (0.80 - attr(private_dat3$sx_admit_rate, "scaled:center"))/(attr(private_dat3$sx_admit_rate, "scaled:scale"))))
```

# Additional Visuals for Presentation

Illustrating the multi-level model
```{r}
public_intslopes$Change_Category

set.seed(20250605)
highlight_set <- slice_sample(public_intslopes, n=2, by = Change_Category)

highlight_data <- highlight_set %>%
  left_join(
    y = public_dat3
    , by = "unit_id"
  ) 

highlight_data %>%
  ggplot(
    aes(x = x_year)
  ) +
  scale_x_continuous(labels = c(2019, 2020, 2021, 2022, 2023)) +
  scale_y_continuous(name = "FTIC Enrollment (log)") +
  # geom_point(
  #   mapping = aes(y = first_time_enrollment, color=Change_Category)
  # ) +
  geom_point(
    mapping = aes(y = log_fyr), size = 3, alpha = 0.33
  ) +
  # geom_line(
  #   mapping = aes(y = log_fyr, group = unit_id, color = Change_Category)
  # ) +
  geom_smooth(
    mapping = aes(y=log_fyr), method="lm", alpha = 0, color = "black", linewidth = 2
  ) +
  scale_color_discrete(guide="none") +
  ggtitle("OLS") +
  theme_classic() +
  theme(
    axis.title.x = element_blank()
  )
ggsave("../Results/glmer_example_0.png", width = 3, height = 3)
```

```{r}
illus_ri <- glmer(first_time_enrollment ~ 1 + x_year + (1 | unit_id), data=highlight_data, family="poisson") # random intercept example
summary(illus_ri)
coef(illus_ri)

ggplot(
  data = highlight_data,
  mapping = aes(x = x_year, y = log_fyr)
) +
  geom_point(size=3, alpha=0.33, mapping=aes(color=factor(unit_id))) +
  geom_abline(
    data = tibble(
      unit_id = rownames(coef(illus_ri)$unit_id) %>% as.factor()
      , `(Intercept)` = coef(illus_ri)$unit_id$`(Intercept)`
      , x_year = coef(illus_ri)$unit_id$x_year
    ), 
    mapping=aes(intercept=`(Intercept)`, slope=x_year, color=factor(unit_id))
  ) +
  scale_x_continuous(labels = c(2019, 2020, 2021, 2022, 2023)) +
  scale_y_continuous(name = "FTIC Enrollment (log)") +
  scale_color_discrete(guide="none") +
  geom_smooth(
    mapping = aes(y=log_fyr), method="lm", alpha = 0, color = "black", linewidth = 2
  ) +
  theme_classic() +
  theme(axis.title.x = element_blank()) +
  ggtitle("Random Intercepts")
ggsave("../Results/glmer_example1.png", width=3, height=3)
```

```{r}
illus_rs <- glmer(first_time_enrollment ~ 1 + x_year + (1 + x_year | unit_id), data=highlight_data, family="poisson") # random slopes example
summary(illus_rs)
coef(illus_rs)

ggplot(
  data = highlight_data,
  mapping = aes(x = x_year, y = log_fyr)
) +
  geom_point(size=3, alpha=0.33, mapping=aes(color=factor(unit_id))) +
  geom_abline(
    data = tibble(
      unit_id = rownames(coef(illus_rs)$unit_id) %>% as.factor()
      , `(Intercept)` = coef(illus_rs)$unit_id$`(Intercept)`
      , x_year = coef(illus_rs)$unit_id$x_year
    ), 
    mapping=aes(intercept=`(Intercept)`, slope=x_year, color=factor(unit_id))
  ) +
  scale_x_continuous(labels = c(2019, 2020, 2021, 2022, 2023)) +
  scale_y_continuous(name = "FTIC Enrollment (log)") +
  scale_color_discrete(guide="none") +
  # geom_abline(
  #   data = tibble(
  #     Global_Intercept = illus_rs@beta[1]
  #     , Global_Slope = illus_rs@beta[2]
  #   )
  #   , mapping = aes(intercept = Global_Intercept, slope=Global_Slope)
  #   , color="black", linewidth=3
  # ) +
  geom_smooth(method="lm", alpha=0, color="black", linewidth=2) +
  theme_classic() +
  theme(axis.title.x = element_blank()) +
  ggtitle("+ Random Slopes")
ggsave("../Results/glmer_example2.png", width=3, height=3)
```

```{r}
ggplot() +
  scale_x_continuous(limits = c(0,4), breaks = c(0,1,2,3,4), labels = c(2019, 2020, 2021, 2022, 2023)) +
  scale_y_continuous(name = "Est. FTIC Enrollment (Log)", limits = c(2,10)) +
  geom_abline(
    data = combined_change %>% mutate(Change_Div = factor(Change_Category, levels = c("Growth (>2.5%)", "Stable (Positive, <2.5%)", "Stable (Negative, <2.5%)", "Decline (>2.5%)"))),
    mapping = aes(intercept = m1_int, slope = m1_slope_time, color=Change_Div), alpha=0.35, linewidth = 2
  ) +
  scale_color_brewer(palette = "RdBu", direction = -1, guide="none") +
  theme_classic()
ggsave("../Results/m1_trend.png", width = 3, height = 3)
```
```{r}
combined_change %>%
  count(sector, Change_Category) %>%
  ungroup() %>%
  group_by(sector) %>%
  mutate(pct = n/sum(n)) %>%
  ungroup() %>%
  mutate(Change_Div = factor(Change_Category, levels = c("Growth (>2.5%)", "Stable (Positive, <2.5%)", "Stable (Negative, <2.5%)", "Decline (>2.5%)"))) %>%
  ggplot(
    aes(x=factor(sector, levels = c("public","private")), y=pct)
  ) +
  geom_bar(stat="identity", mapping=aes(fill=Change_Div), width = 0.33) +
  scale_fill_brewer(palette = "RdBu", direction = -1, name = "Enrollment Trend (Random Slope)") +
  scale_x_discrete(labels = c("Public", "Private")) +
  scale_y_continuous(name = "Percent of Institutions", labels = paste0(c(0,25,50,75,100),"%")) +
  theme_classic() +
  theme(axis.title.x = element_blank())
ggsave("../Results/m1_trend_categories.png", width=5, height=3)
```
```{r}
summary(glm_public_2) ; summary(glm_private_2)
tibble(
  sector = rep(c("Public", "Private"), each = 3),
  parameter = rep(c("Rural vs. Urban", "HS Grads", "Ed. Attain."), times = 2),
  beta = c(
    glm_public_2@beta[3]
    ,glm_public_2@beta[5]
    ,glm_public_2@beta[4]
    ,glm_private_2@beta[3]
    ,glm_private_2@beta[5]
    ,glm_private_2@beta[4]
  ),
  bexp = exp(beta),
  sectorxprm = paste0(sector,"-",parameter) %>% factor(levels = c("Public-Rural vs. Urban","Public-HS Grads","Public-Ed. Attain.","Private-Rural vs. Urban","Private-HS Grads","Private-Ed. Attain."))
) %>%
  ggplot(aes(x=parameter %>% factor(levels = c("Rural vs. Urban", "HS Grads", "Ed. Attain.")), y=bexp)) +
  geom_hline(yintercept = 1, color="#f85c44", linetype=2, linewidth=2) +
  geom_bar(stat="identity", mapping = aes(fill = sectorxprm), position = "dodge") +
  # geom_text(aes(label=round(bexp,2)*100)) +
  scale_fill_manual(
    values = c(
      "#012147"
      ,"#012147"
      ,"#012147"
      ,"#59dce3"
      ,"#59dce3"
      ,"#59dce3"
    ),
    guide = "none"
  ) +
  scale_y_continuous(name = "Rate Ratio (Exp. Beta Coefficient)") +
  scale_x_discrete(labels = c("Rural (vs. Urban)", "HS Grads", "Ed. Attain.")) +
  theme_classic() +
  theme(axis.title.x = element_blank())
ggsave("../Results/m2_exp_beta.png", width = 3, height = 3)

tibble(
  sector = rep(c("Public", "Private"), each = 3) %>% factor(levels = c("Public","Private")),
  parameter = rep(c("Rural vs. Urban", "HS Grads", "Ed. Attain."), times = 2) %>% factor(levels = c("Rural vs. Urban","HS Grads","Ed. Attain.")),
  beta = c(
    glm_public_2@beta[3]
    ,glm_public_2@beta[5]
    ,glm_public_2@beta[4]
    ,glm_private_2@beta[3]
    ,glm_private_2@beta[5]
    ,glm_private_2@beta[4]
  ),
  bexp = exp(beta),
  sectorxprm = paste0(sector,"-",parameter) %>% factor(levels = c("Public-Rural vs. Urban","Public-HS Grads","Public-Ed. Attain.","Private-Rural vs. Urban","Private-HS Grads","Private-Ed. Attain."))
) %>%
  ggplot(mapping = aes(x = sector, y = bexp)) +
  geom_hline(yintercept = 1, linewidth = 1, alpha = 0.33) +
  #geom_point() +
  geom_segment(mapping = aes(y = 1, yend = bexp, color=sector), arrow = arrow(length = unit(0.05, "npc"), type="closed"), linewidth = 1.5) +
  scale_color_manual(values = c("#012147","#59dce3"), guide = "none") +
  scale_y_continuous(name = "Rate Ratio (Exp. Beta Coefficient)", limits = c(0.3,1.3)) +
  facet_wrap(~parameter) +
  theme_classic() +
  theme(axis.title.x = element_blank())
ggsave("../Results/m2_exp_beta_arrows.svg", width = 5, height = 3)

```
```{r}
combined_change %>%
  mutate(sector_f = factor(sector, levels = c("public", "private"), labels = c("Public", "Private"))) %>%
  ggplot(aes(x=m1_slope_time, y=m2_slope_time)) +
  geom_abline(slope = 1, intercept = 0, linewidth = 1.25, color="#bfbfbf") +
  geom_hline(yintercept = 0, color = "#bfbfbf", linewidth = 1.25) +
  geom_point(aes(color=sector_f), size = 3, alpha = 0.65) +
  scale_color_manual(values = c("#012147","#59dce3"), guide = "none") +
  scale_x_continuous(name = "Slope (M1)") +
  scale_y_continuous(name = "Slope (M2)") +
  facet_wrap(~sector_f) +
  theme_classic()
ggsave("../Results/m2_slope_change.svg", width = 4, height = 3)
```
```{r}
summary(glm_public_3) ; summary(glm_private_3)
tibble(
  sector = rep(c("Public", "Private"), each = 5),
  parameter = rep(
    c(
      "Applicants"
      , "Admit Rate"
      , "Non-White %"
      , "Pell %"
      , "Net Price"
    )
    , times = 2
  ) %>% factor(levels = c("Applicants", "Admit Rate", "Non-White %", "Pell %", "Net Price")),
  beta = c(
    glm_public_3@beta[8]
    ,glm_public_3@beta[9]
    ,glm_public_3@beta[6]
    ,glm_public_3@beta[7]
    ,glm_public_3@beta[10]
    ,glm_private_3@beta[8]
    ,glm_private_3@beta[9]
    ,glm_private_3@beta[6]
    ,glm_private_3@beta[7]
    ,glm_private_3@beta[10]
    
  ),
  bexp = exp(beta),
  sectorxprm = paste0(sector,"-",parameter) %>% 
    factor(levels = c(
      "Public-Applicants"
      , "Public-Admit Rate"
      , "Public-Non-White %"
      , "Public-Pell %"
      , "Public-Net Price"
      ,"Private-Applicants"
      , "Private-Admit Rate"
      , "Private-Non-White %"
      , "Private-Pell %"
      , "Private-Net Price"
      )
    )
) %>%
  ggplot(aes(x=parameter, y=bexp)) +
  geom_hline(yintercept = 1, color="#f85c44", linetype=2, linewidth=2) +
  geom_bar(stat="identity", mapping = aes(fill = sectorxprm), position = "dodge") +
  # geom_text(aes(label=round(bexp,2)*100)) +
  scale_fill_manual(
    values = c(
      "#012147"
      ,"#012147"
      ,"#012147"
      ,"#012147"
      ,"#012147"
      ,"#59dce3"
      ,"#59dce3"
      ,"#59dce3"
      ,"#59dce3"
      ,"#59dce3"
    ),
    guide = "none"
  ) +
  scale_y_continuous(name = "Rate Ratio (Exp. Beta Coefficient)") +
  # scale_x_discrete(labels = c("Rural (vs. Urban)", "HS Grads", "Ed. Attain.")) +
  theme_classic() +
  theme(axis.title.x = element_blank())
ggsave("../Results/m3_exp_beta.png", width = 6, height = 3)

tibble(
  sector = rep(c("Public", "Private"), each = 5) %>% factor(levels = c("Public","Private")),
  parameter = rep(
    c(
      "Applicants"
      , "Admit Rate"
      , "Non-White %"
      , "Pell %"
      , "Net Price"
    )
    , times = 2
  ) %>% factor(levels = c("Applicants", "Admit Rate", "Non-White %", "Pell %", "Net Price")),
  beta = c(
    glm_public_3@beta[8]
    ,glm_public_3@beta[9]
    ,glm_public_3@beta[6]
    ,glm_public_3@beta[7]
    ,glm_public_3@beta[10]
    ,glm_private_3@beta[8]
    ,glm_private_3@beta[9]
    ,glm_private_3@beta[6]
    ,glm_private_3@beta[7]
    ,glm_private_3@beta[10]
    
  ),
  bexp = exp(beta),
  sectorxprm = paste0(sector,"-",parameter) %>% 
    factor(levels = c(
      "Public-Applicants"
      , "Public-Admit Rate"
      , "Public-Non-White %"
      , "Public-Pell %"
      , "Public-Net Price"
      ,"Private-Applicants"
      , "Private-Admit Rate"
      , "Private-Non-White %"
      , "Private-Pell %"
      , "Private-Net Price"
      )
    )
) %>%
  ggplot(mapping = aes(x = sector, y = bexp)) +
  geom_hline(yintercept = 1, linewidth = 1, alpha = 0.33) +
  #geom_point() +
  geom_segment(mapping = aes(y = 1, yend = bexp, color=sector), arrow = arrow(length = unit(0.05, "npc"), type="closed"), linewidth = 1.5) +
  scale_color_manual(values = c("#012147","#59dce3"), guide = "none") +
  scale_y_continuous(name = "Rate Ratio (Exp. Beta Coefficient)", limits = c(0.8,1.8)) +
  facet_wrap(~parameter, nrow=1) +
  theme_classic() +
  theme(axis.title.x = element_blank())
ggsave("../Results/m3_exp_beta_arrow.svg", width = 9, height = 3)
```

# Example Visuals

```{r}
set.seed(20250617)
x_ <- tibble(
  Year = 0:4
  , Y = -0.5 * Year + 5 + rnorm(5, mean = 0, sd = 0.5)
)

ggplot(
    x_, aes(x = Year, y = Y)
  ) +
  geom_point(size = 1.5) +
  coord_equal(ylim = c(2,7)) +
  theme_classic() +
  theme(
    axis.title = element_blank()
    , axis.ticks = element_blank()
    , axis.text = element_blank()
  )
ggsave("../Results/regression_ex_0.svg", width=1, height=1)

ggplot(
    x_, aes(x = Year, y = Y)
  ) +
  geom_point(alpha=0.25, size = 1.5) +
  geom_smooth(method="lm", alpha = 0, color="black") +
  coord_equal(ylim = c(2,7)) +
  theme_classic() +
  theme(
    axis.title = element_blank()
    , axis.ticks = element_blank()
    , axis.text = element_blank()
  )
ggsave("../Results/regression_ex_2.svg", width=1, height=1)


```

```{r}
set.seed(20250617)
z_ <- tibble(
  Year = rep(c(0,1,2,3,4),times=3)
  , UnitID = rep(c("A","B","C"),each=5)
  , Slope = case_match(UnitID, "A" ~ 0.5, "B" ~0.25, "C"~-0.6)
  , Intercept = case_match(UnitID, "A" ~ 5, "B" ~ 3.5, "C" ~ 4)
  , Rand = rnorm(15, mean = 0, sd = 0.5)
  , YVAL = Intercept + Slope * Year + Rand
)

zm_ols <- lm(YVAL ~ Year, data = z_)
zm_rin <- lmer(YVAL ~ Year + (1 | UnitID), data = z_)
zm_rsl <- lmer(YVAL ~ Year + (1 + Year | UnitID), data = z_)

ggplot(
  z_, aes(x=Year, y=YVAL)
) +
  geom_point(size = 1.5) +
  geom_abline(intercept = zm_ols$coefficients[1], slope = zm_ols$coefficients[2]) +
  theme_classic() +
  theme(
    axis.title = element_blank()
    , axis.ticks = element_blank()
    , axis.text = element_blank()
  )
ggsave("../Results/multilevel_ex_0.svg", width=1, height=1)

ggplot(
  z_, aes(x=Year, y=YVAL, color=UnitID)
) +
  geom_point(size=1.5) +
  geom_abline(
    data = tibble(
      UnitID = c("A","B","C")
      , MINT = coef(zm_rin)$UnitID[,1]
      , MSLP = coef(zm_rin)$UnitID[,2]
    )
    , mapping = aes(intercept = MINT, slope = MSLP, color=UnitID)
  ) +
  theme_classic() +
  theme(
    axis.title = element_blank()
    , axis.ticks = element_blank()
    , axis.text = element_blank()
    , legend.position = "none"
  )
ggsave("../Results/multilevel_ex_1.svg", width=1, height=1)


ggplot(
  z_, aes(x=Year, y=YVAL, color=UnitID)
) +
  geom_point(size=1.5) +
  geom_abline(
    data = tibble(
      UnitID = c("A","B","C")
      , MINT = coef(zm_rsl)$UnitID[,1]
      , MSLP = coef(zm_rsl)$UnitID[,2]
    )
    , mapping = aes(intercept = MINT, slope = MSLP, color=UnitID)
  ) +
  theme_classic() +
  theme(
    axis.title = element_blank()
    , axis.ticks = element_blank()
    , axis.text = element_blank()
    , legend.position = "none"
  )
ggsave("../Results/multilevel_ex_2.svg", width=1, height=1)


```


```{r}
set.seed(20250605)
highlight_set <- slice_sample(public_intslopes, n=2, by = Change_Category)

highlight_data <- highlight_set %>%
  left_join(
    y = public_dat3
    , by = "unit_id"
  ) 

illus_ols <- lm(log_fyr ~ x_year, data = highlight_set)
highlight_data %>%
  ggplot(
    aes(x = x_year)
  ) +
  scale_x_continuous(labels = c(2019, 2020, 2021, 2022, 2023)) +
  scale_y_continuous(name = "FTIC Enrollment (log)") +
  # geom_point(
  #   mapping = aes(y = first_time_enrollment, color=Change_Category)
  # ) +
  geom_point(
    mapping = aes(y = log_fyr), size = 3, alpha = 0.33
  ) +
  # geom_line(
  #   mapping = aes(y = log_fyr, group = unit_id, color = Change_Category)
  # ) +
  geom_smooth(
    mapping = aes(y=log_fyr), method="lm", alpha = 0, color = "black", linewidth = 2
  ) +
  scale_color_discrete(guide="none") +
  ggtitle("OLS") +
  theme_classic() +
  theme(
    axis.title.x = element_blank()
  )
ggsave("../Results/glmer_example_0.png", width = 3, height = 3)

illus_ri <- glmer(first_time_enrollment ~ 1 + x_year + (1 | unit_id), data=highlight_data, family="poisson") # random intercept example
summary(illus_ri)
coef(illus_ri)

ggplot(
  data = highlight_data,
  mapping = aes(x = x_year, y = log_fyr)
) +
  geom_point(size=3, alpha=0.33, mapping=aes(color=factor(unit_id))) +
  geom_abline(
    data = tibble(
      unit_id = rownames(coef(illus_ri)$unit_id) %>% as.factor()
      , `(Intercept)` = coef(illus_ri)$unit_id$`(Intercept)`
      , x_year = coef(illus_ri)$unit_id$x_year
    ), 
    mapping=aes(intercept=`(Intercept)`, slope=x_year, color=factor(unit_id))
  ) +
  scale_x_continuous(labels = c(2019, 2020, 2021, 2022, 2023)) +
  scale_y_continuous(name = "FTIC Enrollment (log)") +
  scale_color_discrete(guide="none") +
  geom_smooth(
    mapping = aes(y=log_fyr), method="lm", alpha = 0, color = "black", linewidth = 2
  ) +
  theme_classic() +
  theme(axis.title.x = element_blank()) +
  ggtitle("Random Intercepts")
ggsave("../Results/glmer_example1.png", width=3, height=3)

illus_rs <- glmer(first_time_enrollment ~ 1 + x_year + (1 + x_year | unit_id), data=highlight_data, family="poisson") # random slopes example
summary(illus_rs)
coef(illus_rs)

ggplot(
  data = highlight_data,
  mapping = aes(x = x_year, y = log_fyr)
) +
  geom_point(size=3, alpha=0.33, mapping=aes(color=factor(unit_id))) +
  geom_abline(
    data = tibble(
      unit_id = rownames(coef(illus_rs)$unit_id) %>% as.factor()
      , `(Intercept)` = coef(illus_rs)$unit_id$`(Intercept)`
      , x_year = coef(illus_rs)$unit_id$x_year
    ), 
    mapping=aes(intercept=`(Intercept)`, slope=x_year, color=factor(unit_id))
  ) +
  scale_x_continuous(labels = c(2019, 2020, 2021, 2022, 2023)) +
  scale_y_continuous(name = "FTIC Enrollment (log)") +
  scale_color_discrete(guide="none") +
  # geom_abline(
  #   data = tibble(
  #     Global_Intercept = illus_rs@beta[1]
  #     , Global_Slope = illus_rs@beta[2]
  #   )
  #   , mapping = aes(intercept = Global_Intercept, slope=Global_Slope)
  #   , color="black", linewidth=3
  # ) +
  geom_smooth(method="lm", alpha=0, color="black", linewidth=2) +
  theme_classic() +
  theme(axis.title.x = element_blank()) +
  ggtitle("+ Random Slopes")
ggsave("../Results/glmer_example2.png", width=3, height=3)
```

