---
title: "Multi-Level Modeling of Enrollment"
output: html_notebook
---

# Setup

```{r}
library(tidyverse)
library(lme4)
```
```{r}
full_dat3 <- read_csv("../Data/full_dat3.csv")
str(full_dat3)
```

```{r}
public_dat3 <- full_dat3 %>% filter(sector == "public")
private_dat3 <- full_dat3 %>% filter(sector == "private")
```

```{r}
public_dat3 %>%
  group_by(unit_id) %>%
  summarise(N_year = length(unique(year))) %>%
  ungroup() %>%
  count(N_year)
  
private_dat3 %>%
  group_by(unit_id) %>%
  summarise(N_year = length(unique(year))) %>%
  ungroup() %>%
  count(N_year)
```
36 out of 530 public institutions do not have 5 years of IPEDS data
77 out of 1076 private institutions do not have 5 years of IPEDS data

```{r}
public_dat3 %>%
  group_by(unit_id) %>%
  summarise(Min_Enrl = min(overall_enrollment), Max_Enrl = max(overall_enrollment)) %>%
  ungroup() %>%
  count(Min_Enrl < 1000, Max_Enrl > 1000)
  # filter(Min_Enrl < 1000, Max_Enrl > 1000)
private_dat3 %>%
  group_by(unit_id) %>%
  summarise(Min_Enrl = min(overall_enrollment), Max_Enrl = max(overall_enrollment)) %>%
  ungroup() %>%
  count(Min_Enrl < 1000, Max_Enrl > 1000)
  # filter(Min_Enrl < 1000, Max_Enrl > 1000)
```
2 out of 530 public institutions cross the 1000 overall enrollment threshold in this period.
90 out of 1076 private institutions cross the 1000 overall enrollment threshold in this period.

```{r}
public_dat3 %>%
  group_by(unit_id) %>%
  filter(year == min(year)) %>%
  ungroup() %>%
  filter(overall_enrollment >= 1000) %>%
  distinct(unit_id) %>%
  left_join(
    y = public_dat3,
    by = "unit_id"
  )

private_dat3 %>%
  group_by (unit_id) %>%
  filter(year == min(year)) %>%
  ungroup() %>%
  filter(overall_enrollment >= 1000) %>%
  distinct(unit_id) %>%
  left_join(
    y = private_dat3,
    by = "unit_id"
  )

private_dat3 %>%
  group_by (unit_id) %>%
  filter(year == min(year)) %>%
  ungroup() %>%
  filter(overall_enrollment < 1000) %>%
  mutate(
    Enrl_Level = case_when(
      overall_enrollment >= 800 ~ "800+",
      overall_enrollment >= 600 ~ "600-800",
      overall_enrollment >= 400 ~ "400-600",
      overall_enrollment >= 200 ~ "200-400",
      overall_enrollment > 0 ~ "1-200"
    )
  ) %>%
  count(Enrl_Level)

```
504 of 530 public institutions enrolled at least 1000 students in their first reporting year in this period.
679 of 1076 private institutions enrolled at least 1000 students in their first reporting year in this period.

```{r}
public_dat3 %>%
  distinct(unit_id, carnegie) %>%
  count(carnegie)
private_dat3 %>%
  distinct(unit_id, carnegie) %>%
  count(carnegie)

table(full_dat3$carnegie)
table(private_dat3_x5$carnegie)
```
26 out of 530 public institutions were not in the Carnegie universe.
15 out of 1076 private institutions were not in the Carnegie universe.

```{r}
public_dat3_x <- public_dat3 %>%
  # In the Carnegie universe
  filter(carnegie != "Not Carnegie") %>%
  # At least 1000 overall enrollment in first reporting year
  group_by(unit_id) %>%
  filter(year == min(year)) %>%
  ungroup() %>%
  filter(overall_enrollment >= 1000) %>%
  distinct(unit_id) %>%
  left_join(
    y = public_dat3,
    by = "unit_id"
  ) %>%
  # Has all 5 years of data
  group_by(unit_id) %>%
  summarise(N_year = length(unique(year))) %>%
  ungroup() %>%
  left_join(
    y = public_dat3,
    by = "unit_id"
  ) %>%
  mutate(
    year_0 = year - min(year),
    # Log transforms: first-time enrollment; HS grads by state; applicants; average net price with aid
    log_fyr = log(first_time_enrollment),
    log_hs = log(hs_grad_state),
    log_apps = log(applicants),
    log_price = log(avg_net_price_w_aid),
    sy_fyr = scale(log_fyr),
    x_year = year_0,
    sx_hs = scale(log_hs), # log this
    sx_bach = scale(pct_bachelors_2018_state),
    x_rural = factor(locale_rural),
    sx_pell = scale(pct_pell),
    sx_race_minority = scale(pct_race_minority),
    sx_price = scale(log_price), # don't log this
    sx_apps = scale(log_apps), # log this
    sx_admit_rate = scale(admissions_rate)
  )

public_dat3_x5 <- filter(public_dat3_x, N_year == 5)

private_dat3_x <- private_dat3 %>%
  # In the Carnegie universe
  filter(carnegie != "Not Carnegie") %>%
  # At least 1000 overall enrollment in first reporting year
  group_by(unit_id) %>%
  filter(year == min(year)) %>%
  ungroup() %>%
  filter(overall_enrollment >= 1000) %>%
  distinct(unit_id) %>%
  left_join(
    y = private_dat3,
    by = "unit_id"
  ) %>%
  # Has all 5 years of data
  group_by(unit_id) %>%
  summarise(N_year = length(unique(year))) %>%
  ungroup() %>%
  left_join(
    y = private_dat3,
    by = "unit_id"
  ) %>%
  mutate(
    year_0 = year - min(year),
    # Log transforms: first-time enrollment; HS grads by state; applicants; average net price with aid
    log_fyr = log(first_time_enrollment),
    log_hs = log(hs_grad_state),
    log_apps = log(applicants),
    log_price = log(avg_net_price_w_aid),
    sy_fyr = scale(log_fyr),
    x_year = year_0,
    sx_hs = scale(log_hs),
    sx_bach = scale(pct_bachelors_2018_state),
    x_rural = factor(locale_rural),
    sx_pell = scale(pct_pell),
    sx_race_minority = scale(pct_race_minority),
    sx_price = scale(log_price),
    sx_apps = scale(log_apps),
    sx_admit_rate = scale(admissions_rate)
  )

private_dat3_x5 <- filter(private_dat3_x, N_year == 5)
```

```{r}
distinct(public_dat3_x5, unit_id) ; distinct(private_dat3_x5, unit_id)
distinct(public_dat3_x, unit_id) ; distinct(private_dat3_x, unit_id)

```
479 out of 530 public institutions meet the criteria ...
662 out of 1076 private institutions meet the criteria ...

# Model 1: Basic Growth Model (Null Model)

```{r}
formula_1 <- formula(
  first_time_enrollment ~
    1
    + x_year
    # + (1 + x_year | carnegie)
    + (1 + x_year | unit_id)
)
```

## Model Estimation: Public Institutions

```{r}
m_public_1 <- lme4::lmer(
  formula = formula_1,
  data = public_dat3_x5
)
summary(m_public_1)
coef(m_public_1)
# fitted(m_public_1)
# resid(m_public_1)
```

## Model Estimation: Private Institutions

```{r}
m_private_1 <- lme4::lmer(
  formula = formula_1,
  data = private_dat3_x5
)
summary(m_private_1)
coef(m_private_1)
# fitted(m_private_1)
# resid(m_private_1)
```

## Results Visualization: Overall FYR recruitment trends

```{r}
public_fyr_normalized <- public_dat3_x5 %>%
  filter(x_year==0) %>%
  select(unit_id, first_time_enrollment, overall_enrollment) %>%
  rename(baseline_enrollment = first_time_enrollment, baseline_overall=overall_enrollment) %>%
  left_join(
    y = public_dat3_x5 %>% select(unit_id, institution_name, carnegie, year, x_year, first_time_enrollment),
    by = "unit_id"
  ) %>%
  mutate(
    normalized_enrollment = first_time_enrollment / baseline_enrollment,
    size_category = case_when(
      baseline_overall >= 1000 & baseline_overall < 5000 ~ "1k-5k",
      baseline_overall >= 5000 & baseline_overall < 10000 ~ "5k-10k",
      baseline_overall >= 10000 & baseline_overall < 20000 ~ "10k-20k",
      baseline_overall >= 20000 ~ "20k+"
    ) %>% factor(levels = c("1k-5k","5k-10k","10k-20k","20k+")),
    carnegie_category = factor(
      carnegie,
      levels = c("R1", "R2", "R0", "M", "Baccalaureate", "Special or Tribal"),
      labels = c("R1", "R2", "D/PU", "M", "B", "SF&TCU")
    )
  ) %>%
  left_join(
    y = tibble(
      unit_id = rownames(coef(m_public_1)$unit_id) %>% as.numeric(),
      time_slope = coef(m_public_1)$unit_id[,"x_year"],
      enrl_intcpt = coef(m_public_1)$unit_id[,"(Intercept)"],
      growth_pct = time_slope / enrl_intcpt,
      growth_bucket = case_when(
        growth_pct < -0.05 ~ "Linear Decline more than 5% per year",
        growth_pct < -0.01 ~ "Linear Decline 1 to 5% per year",
        growth_pct <= 0.01 ~ "Linear Change less than 1% per year",
        growth_pct <= 0.05 ~ "Linear Growth 1 to 5% per year",
        growth_pct > 0.05 ~ "Linear Growth more than 5% per year"
      ) %>% factor(levels = c("Linear Decline more than 5% per year","Linear Decline 1 to 5% per year","Linear Change less than 1% per year","Linear Growth 1 to 5% per year","Linear Growth more than 5% per year"))
    ) ,
    by = "unit_id"
  )

public_fyr_slope_avg <- public_fyr_normalized %>%
  group_by(size_category, carnegie_category) %>%
  summarise(Avg_Slope = mean(growth_pct))
```

```{r}
# public_fyr_normalized %>%
#   ggplot(aes(x=x_year, y=normalized_enrollment, group=unit_id, color=growth_bucket)) +
#   geom_line(alpha = 0.9) +
#   scale_color_brewer(type="div", palette = "RdYlBu") +
#   facet_wrap(~size_category) +
#   theme_classic()
# 
# public_fyr_normalized %>%
#   ggplot(aes(x=x_year, y=normalized_enrollment, group=unit_id, color=growth_bucket)) +
#   geom_line(alpha = 0.9) +
#   scale_color_brewer(type="div", palette = "RdYlBu") +
#   facet_wrap(~carnegie_category) +
#   theme_classic()

public_fyr_normalized %>%
  ggplot(aes(x=x_year, y=normalized_enrollment)) +
  geom_line(mapping=aes(group=unit_id, color=growth_bucket), alpha = 0.9) +
  scale_color_brewer(type="div", palette = "RdYlBu") +
  geom_text(data=public_fyr_slope_avg, mapping=aes(x=0.5,y=3,label=round(Avg_Slope, 3))) +
  scale_y_continuous(name = "Normalized FYR Recruitment") +
  scale_x_continuous(name = "", labels = c(2019, 2020, 2021, 2022, 2023)) +
  facet_grid(size_category~carnegie_category) +
  # geom_abline(data=public_fyr_slope_avg, mapping=aes(slope=Avg_Slope, intercept=1), linewidth=1, linetype = 1) +
  theme_classic()

```

```{r}
private_fyr_normalized <- private_dat3_x5 %>%
  filter(x_year==0) %>%
  select(unit_id, first_time_enrollment, overall_enrollment) %>%
  rename(baseline_enrollment = first_time_enrollment, baseline_overall=overall_enrollment) %>%
  left_join(
    y = private_dat3_x5 %>% select(unit_id, institution_name, carnegie, year, x_year, first_time_enrollment),
    by = "unit_id"
  ) %>%
  mutate(
    normalized_enrollment = first_time_enrollment / baseline_enrollment,
    size_category = case_when(
      baseline_overall >= 1000 & baseline_overall < 5000 ~ "1k-5k",
      baseline_overall >= 5000 & baseline_overall < 10000 ~ "5k-10k",
      baseline_overall >= 10000 & baseline_overall < 20000 ~ "10k-20k",
      baseline_overall >= 20000 ~ "20k+"
    ) %>% factor(levels = c("1k-5k","5k-10k","10k-20k","20k+")),
    carnegie_category = factor(
      carnegie,
      levels = c("R1", "R2", "R0", "M", "Baccalaureate", "Special or Tribal"),
      labels = c("R1", "R2", "D/PU", "M", "B", "SF&TCU")
    )
  ) %>%
  left_join(
    y = tibble(
      unit_id = rownames(coef(m_private_1)$unit_id) %>% as.numeric(),
      time_slope = coef(m_private_1)$unit_id[,"x_year"],
      enrl_intcpt = coef(m_private_1)$unit_id[,"(Intercept)"],
      growth_pct = time_slope / enrl_intcpt,
      growth_bucket = case_when(
        growth_pct < -0.05 ~ "Linear Decline more than 5% per year",
        growth_pct < -0.01 ~ "Linear Decline 1 to 5% per year",
        growth_pct <= 0.01 ~ "Linear Change less than 1% per year",
        growth_pct <= 0.05 ~ "Linear Growth 1 to 5% per year",
        growth_pct > 0.05 ~ "Linear Growth more than 5% per year"
      ) %>% factor(levels = c("Linear Decline more than 5% per year","Linear Decline 1 to 5% per year","Linear Change less than 1% per year","Linear Growth 1 to 5% per year","Linear Growth more than 5% per year"))
    ) ,
    by = "unit_id"
  )

private_fyr_slope_avg <- private_fyr_normalized %>%
  group_by(size_category, carnegie_category) %>%
  summarise(Avg_Slope = mean(growth_pct))
```

```{r}
private_fyr_normalized %>%
  ggplot(aes(x=x_year, y=normalized_enrollment)) +
  geom_line(mapping=aes(group=unit_id, color=growth_bucket), alpha = 0.9) +
  scale_color_brewer(type="div", palette = "RdYlBu") +
  geom_text(data=private_fyr_slope_avg, mapping=aes(x=0.5,y=3,label=round(Avg_Slope, 3))) +
  scale_y_continuous(name = "Normalized FYR Recruitment") +
  scale_x_continuous(name = "", labels = c(2019, 2020, 2021, 2022, 2023)) +
  facet_grid(size_category~carnegie_category) +
  # geom_abline(data=private_fyr_slope_avg, mapping=aes(slope=Avg_Slope, intercept=1), linewidth=1, linetype = 1) +
  theme_classic()

```

```{r}
public_fyr_normalized %>%
  distinct(unit_id, growth_pct, growth_bucket) %>%
  count(growth_bucket) %>%
  mutate(pct = round(n/sum(n), 2))
```



