---
title: "Multi-Level Modeling of Enrollment"
output: html_notebook
---

# Setup

```{r}
library(tidyverse)
library(lme4)
```
```{r}
full_dat3 <- read_csv("../Data/full_dat3.csv")
str(full_dat3)
```

```{r}
public_dat3 <- full_dat3 %>% filter(sector == "public")
private_dat3 <- full_dat3 %>% filter(sector == "private")
```

```{r}
public_dat3 %>%
  group_by(unit_id) %>%
  summarise(N_year = length(unique(year))) %>%
  ungroup() %>%
  count(N_year)
  
private_dat3 %>%
  group_by(unit_id) %>%
  summarise(N_year = length(unique(year))) %>%
  ungroup() %>%
  count(N_year)
```
36 out of 530 public institutions do not have 5 years of IPEDS data
77 out of 1076 private institutions do not have 5 years of IPEDS data

```{r}
public_dat3 %>%
  group_by(unit_id) %>%
  summarise(Min_Enrl = min(overall_enrollment), Max_Enrl = max(overall_enrollment)) %>%
  ungroup() %>%
  count(Min_Enrl < 1000, Max_Enrl > 1000)
  # filter(Min_Enrl < 1000, Max_Enrl > 1000)
private_dat3 %>%
  group_by(unit_id) %>%
  summarise(Min_Enrl = min(overall_enrollment), Max_Enrl = max(overall_enrollment)) %>%
  ungroup() %>%
  count(Min_Enrl < 1000, Max_Enrl > 1000)
  # filter(Min_Enrl < 1000, Max_Enrl > 1000)
```
2 out of 530 public institutions cross the 1000 overall enrollment threshold in this period.
90 out of 1076 private institutions cross the 1000 overall enrollment threshold in this period.

```{r}
public_dat3 %>%
  group_by(unit_id) %>%
  filter(year == min(year)) %>%
  ungroup() %>%
  filter(overall_enrollment >= 1000) %>%
  distinct(unit_id) %>%
  left_join(
    y = public_dat3,
    by = "unit_id"
  )

private_dat3 %>%
  group_by (unit_id) %>%
  filter(year == min(year)) %>%
  ungroup() %>%
  filter(overall_enrollment >= 1000) %>%
  distinct(unit_id) %>%
  left_join(
    y = private_dat3,
    by = "unit_id"
  )

private_dat3 %>%
  group_by (unit_id) %>%
  filter(year == min(year)) %>%
  ungroup() %>%
  filter(overall_enrollment < 1000) %>%
  mutate(
    Enrl_Level = case_when(
      overall_enrollment >= 800 ~ "800+",
      overall_enrollment >= 600 ~ "600-800",
      overall_enrollment >= 400 ~ "400-600",
      overall_enrollment >= 200 ~ "200-400",
      overall_enrollment > 0 ~ "1-200"
    )
  ) %>%
  count(Enrl_Level)

```
504 of 530 public institutions enrolled at least 1000 students in their first reporting year in this period.
679 of 1076 private institutions enrolled at least 1000 students in their first reporting year in this period.

```{r}
public_dat3 %>%
  distinct(unit_id, carnegie) %>%
  count(carnegie)
private_dat3 %>%
  distinct(unit_id, carnegie) %>%
  count(carnegie)

# table(full_dat3$carnegie)
# table(private_dat3_x5$carnegie)
```
26 out of 530 public institutions were not in the Carnegie universe.
15 out of 1076 private institutions were not in the Carnegie universe.

```{r}
public_dat3_x <- public_dat3 %>%
  # In the Carnegie universe
  filter(carnegie != "Not Carnegie") %>%
  # At least 1000 overall enrollment in first reporting year
  group_by(unit_id) %>%
  filter(year == min(year)) %>%
  ungroup() %>%
  filter(overall_enrollment >= 1000) %>%
  distinct(unit_id) %>%
  left_join(
    y = public_dat3,
    by = "unit_id"
  ) %>%
  # Has all 5 years of data
  group_by(unit_id) %>%
  summarise(N_year = length(unique(year))) %>%
  ungroup() %>%
  left_join(
    y = public_dat3,
    by = "unit_id"
  ) %>%
  mutate(
    year_0 = year - min(year),
    # Log transforms: first-time enrollment; HS grads by state; applicants; average net price with aid
    log_fyr = log(first_time_enrollment),
    log_hs = log(hs_grad_state),
    log_apps = log(applicants),
    log_price = log(avg_net_price_w_aid),
    sy_fyr = scale(log_fyr),
    x_year = year_0,
    sx_hs = scale(log_hs), # log this
    sx_bach = scale(pct_bachelors_2018_state),
    x_rural = factor(locale_rural),
    sx_pell = scale(pct_pell),
    sx_race_minority = scale(pct_race_minority),
    sx_price = scale(log_price), # don't log this
    sx_apps = scale(log_apps), # log this
    sx_admit_rate = scale(admissions_rate)
  )

public_dat3_x5 <- filter(public_dat3_x, N_year == 5)

private_dat3_x <- private_dat3 %>%
  # In the Carnegie universe
  filter(carnegie != "Not Carnegie") %>%
  # At least 1000 overall enrollment in first reporting year
  group_by(unit_id) %>%
  filter(year == min(year)) %>%
  ungroup() %>%
  filter(overall_enrollment >= 1000) %>%
  distinct(unit_id) %>%
  left_join(
    y = private_dat3,
    by = "unit_id"
  ) %>%
  # Has all 5 years of data
  group_by(unit_id) %>%
  summarise(N_year = length(unique(year))) %>%
  ungroup() %>%
  left_join(
    y = private_dat3,
    by = "unit_id"
  ) %>%
  mutate(
    year_0 = year - min(year),
    # Log transforms: first-time enrollment; HS grads by state; applicants; average net price with aid
    log_fyr = log(first_time_enrollment),
    log_hs = log(hs_grad_state),
    log_apps = log(applicants),
    log_price = log(avg_net_price_w_aid),
    sy_fyr = scale(log_fyr),
    x_year = year_0,
    sx_hs = scale(log_hs),
    sx_bach = scale(pct_bachelors_2018_state),
    x_rural = factor(locale_rural),
    sx_pell = scale(pct_pell),
    sx_race_minority = scale(pct_race_minority),
    sx_price = scale(log_price),
    sx_apps = scale(log_apps),
    sx_admit_rate = scale(admissions_rate)
  )

private_dat3_x5 <- filter(private_dat3_x, N_year == 5)
```

```{r}
distinct(public_dat3_x5, unit_id) ; distinct(private_dat3_x5, unit_id)
distinct(public_dat3_x, unit_id) ; distinct(private_dat3_x, unit_id)

```
479 out of 530 public institutions meet the criteria ...
662 out of 1076 private institutions meet the criteria ...

# Model 1: Basic Growth Model (Null Model)

```{r}
formula_1 <- formula(
  first_time_enrollment ~
    1
    + x_year
    # + (1 + x_year | carnegie)
    + (1 + x_year | unit_id)
)
```

## Model Estimation: Public Institutions

```{r}
m_public_1 <- lme4::lmer(
  formula = formula_1,
  data = public_dat3_x5
)
summary(m_public_1)
coef(m_public_1)
# fitted(m_public_1)
# resid(m_public_1)
```

## Model Estimation: Private Institutions

```{r}
m_private_1 <- lme4::lmer(
  formula = formula_1,
  data = private_dat3_x5
)
summary(m_private_1)
coef(m_private_1)
# fitted(m_private_1)
# resid(m_private_1)
```

## Results Visualization: Overall FYR recruitment trends

```{r}
public_fyr_normalized <- public_dat3_x5 %>%
  filter(x_year==0) %>%
  select(unit_id, first_time_enrollment, overall_enrollment) %>%
  rename(baseline_enrollment = first_time_enrollment, baseline_overall=overall_enrollment) %>%
  left_join(
    y = public_dat3_x5 %>% select(unit_id, institution_name, carnegie, year, x_year, first_time_enrollment),
    by = "unit_id"
  ) %>%
  mutate(
    normalized_enrollment = first_time_enrollment / baseline_enrollment,
    size_category = case_when(
      baseline_overall >= 1000 & baseline_overall < 5000 ~ "1k-5k",
      baseline_overall >= 5000 & baseline_overall < 10000 ~ "5k-10k",
      baseline_overall >= 10000 & baseline_overall < 20000 ~ "10k-20k",
      baseline_overall >= 20000 ~ "20k+"
    ) %>% factor(levels = c("1k-5k","5k-10k","10k-20k","20k+")),
    carnegie_category = factor(
      carnegie,
      levels = c("R1", "R2", "R0", "M", "Baccalaureate", "Special or Tribal"),
      labels = c("R1", "R2", "D/PU", "M", "B", "SF&TCU")
    )
  ) %>%
  left_join(
    y = tibble(
      unit_id = rownames(coef(m_public_1)$unit_id) %>% as.numeric(),
      time_slope = coef(m_public_1)$unit_id[,"x_year"],
      enrl_intcpt = coef(m_public_1)$unit_id[,"(Intercept)"],
      growth_pct = time_slope / enrl_intcpt,
      growth_bucket = case_when(
        growth_pct < -0.05 ~ "Linear Decline more than 5% per year",
        growth_pct < -0.01 ~ "Linear Decline 1 to 5% per year",
        growth_pct <= 0.01 ~ "Linear Change less than 1% per year",
        growth_pct <= 0.05 ~ "Linear Growth 1 to 5% per year",
        growth_pct > 0.05 ~ "Linear Growth more than 5% per year"
      ) %>% factor(levels = c("Linear Decline more than 5% per year","Linear Decline 1 to 5% per year","Linear Change less than 1% per year","Linear Growth 1 to 5% per year","Linear Growth more than 5% per year"))
    ) ,
    by = "unit_id"
  )

public_fyr_slope_avg <- public_fyr_normalized %>%
  group_by(size_category, carnegie_category) %>%
  summarise(Avg_Slope = mean(growth_pct))
```

```{r}
# public_fyr_normalized %>%
#   ggplot(aes(x=x_year, y=normalized_enrollment, group=unit_id, color=growth_bucket)) +
#   geom_line(alpha = 0.9) +
#   scale_color_brewer(type="div", palette = "RdYlBu") +
#   facet_wrap(~size_category) +
#   theme_classic()
# 
# public_fyr_normalized %>%
#   ggplot(aes(x=x_year, y=normalized_enrollment, group=unit_id, color=growth_bucket)) +
#   geom_line(alpha = 0.9) +
#   scale_color_brewer(type="div", palette = "RdYlBu") +
#   facet_wrap(~carnegie_category) +
#   theme_classic()

public_fyr_normalized %>%
  ggplot(aes(x=x_year, y=normalized_enrollment)) +
  geom_line(mapping=aes(group=unit_id, color=growth_bucket), alpha = 0.9) +
  scale_color_brewer(type="div", palette = "RdYlBu") +
  geom_text(data=public_fyr_slope_avg, mapping=aes(x=0.5,y=3,label=round(Avg_Slope, 3))) +
  scale_y_continuous(name = "Normalized FYR Recruitment") +
  scale_x_continuous(name = "", labels = c(2019, 2020, 2021, 2022, 2023)) +
  # facet_grid(size_category~carnegie_category) +
  # geom_abline(data=public_fyr_slope_avg, mapping=aes(slope=Avg_Slope, intercept=1), linewidth=1, linetype = 1) +
  theme_classic() +
  theme(legend.position = "bottom")
ggsave("../Results/normalized_fyr_public.png")

```

```{r}
private_fyr_normalized <- private_dat3_x5 %>%
  filter(x_year==0) %>%
  select(unit_id, first_time_enrollment, overall_enrollment) %>%
  rename(baseline_enrollment = first_time_enrollment, baseline_overall=overall_enrollment) %>%
  left_join(
    y = private_dat3_x5 %>% select(unit_id, institution_name, carnegie, year, x_year, first_time_enrollment),
    by = "unit_id"
  ) %>%
  mutate(
    normalized_enrollment = first_time_enrollment / baseline_enrollment,
    size_category = case_when(
      baseline_overall >= 1000 & baseline_overall < 5000 ~ "1k-5k",
      baseline_overall >= 5000 & baseline_overall < 10000 ~ "5k-10k",
      baseline_overall >= 10000 & baseline_overall < 20000 ~ "10k-20k",
      baseline_overall >= 20000 ~ "20k+"
    ) %>% factor(levels = c("1k-5k","5k-10k","10k-20k","20k+")),
    carnegie_category = factor(
      carnegie,
      levels = c("R1", "R2", "R0", "M", "Baccalaureate", "Special or Tribal"),
      labels = c("R1", "R2", "D/PU", "M", "B", "SF&TCU")
    )
  ) %>%
  left_join(
    y = tibble(
      unit_id = rownames(coef(m_private_1)$unit_id) %>% as.numeric(),
      time_slope = coef(m_private_1)$unit_id[,"x_year"],
      enrl_intcpt = coef(m_private_1)$unit_id[,"(Intercept)"],
      growth_pct = time_slope / enrl_intcpt,
      growth_bucket = case_when(
        growth_pct < -0.05 ~ "Linear Decline more than 5% per year",
        growth_pct < -0.01 ~ "Linear Decline 1 to 5% per year",
        growth_pct <= 0.01 ~ "Linear Change less than 1% per year",
        growth_pct <= 0.05 ~ "Linear Growth 1 to 5% per year",
        growth_pct > 0.05 ~ "Linear Growth more than 5% per year"
      ) %>% factor(levels = c("Linear Decline more than 5% per year","Linear Decline 1 to 5% per year","Linear Change less than 1% per year","Linear Growth 1 to 5% per year","Linear Growth more than 5% per year"))
    ) ,
    by = "unit_id"
  )

private_fyr_slope_avg <- private_fyr_normalized %>%
  group_by(size_category, carnegie_category) %>%
  summarise(Avg_Slope = mean(growth_pct))
```

```{r}
private_fyr_normalized %>%
  ggplot(aes(x=x_year, y=normalized_enrollment)) +
  geom_line(mapping=aes(group=unit_id, color=growth_bucket), alpha = 0.9) +
  scale_color_brewer(type="div", palette = "RdYlBu") +
  geom_text(data=private_fyr_slope_avg, mapping=aes(x=0.5,y=3,label=round(Avg_Slope, 3))) +
  scale_y_continuous(name = "Normalized FYR Recruitment") +
  scale_x_continuous(name = "", labels = c(2019, 2020, 2021, 2022, 2023)) +
  facet_grid(size_category~carnegie_category) +
  # geom_abline(data=private_fyr_slope_avg, mapping=aes(slope=Avg_Slope, intercept=1), linewidth=1, linetype = 1) +
  theme_classic() +
  theme(legend.position = "bottom")
ggsave("../Results/normalized_fyr_private.png")

```

```{r}
public_fyr_normalized %>%
  distinct(unit_id, growth_pct, growth_bucket) %>%
  count(growth_bucket) %>%
  mutate(pct = round(n/sum(n), 2))

private_fyr_normalized %>%
  distinct(unit_id, growth_pct, growth_bucket) %>%
  count(growth_bucket) %>%
  mutate(pct = round(n/sum(n), 2))

public_fyr_normalized %>%
  distinct(unit_id, growth_pct, growth_bucket, carnegie, size_category) %>%
  group_by(carnegie, size_category) %>%
  summarise(
    tot_decline = sum(growth_bucket %in% c("Linear Decline 1 to 5% per year","Linear Decline more than 5% per year")),
    tot_inst = length(unique(unit_id)),
    decline_frac = paste0(tot_decline,"/",tot_inst, " (", 100*round(tot_decline/tot_inst,2), "%)")
  )

private_fyr_normalized %>%
  distinct(unit_id, growth_pct, growth_bucket, carnegie, size_category) %>%
  group_by(carnegie, size_category) %>%
  summarise(
    tot_decline = sum(growth_bucket %in% c("Linear Decline 1 to 5% per year","Linear Decline more than 5% per year")),
    tot_inst = length(unique(unit_id)),
    decline_frac = paste0(tot_decline,"/",tot_inst, " (", 100*round(tot_decline/tot_inst,2), "%)")
  )

```

# Model 2: Geographic and Demographic Context

```{r}
formula_2 <- formula(
  first_time_enrollment ~
    1
    + x_year
    + x_rural
    + sx_bach # Percent of adults with bachelors degree in the state where the institution is located, standardized
    + sx_hs # The number of high school graduates that year, logged and standardized
    # + (1 + x_year | carnegie)
    + (1 + x_year | unit_id)
)
```

## Model Estimation: Public

```{r}
m_public_2 <- lmer(formula_2, public_dat3_x5)
summary(m_public_2)
```
```{r}
coef(m_public_1)$unit_id
coef(m_public_2)$unit_id

tibble(
  unit_id = rownames(coef(m_public_1)$unit_id) %>% as.numeric(),
  M1_Intcpt = coef(m_public_1)$unit_id[,"(Intercept)"],
  M1_Slope = coef(m_public_1)$unit_id[,"x_year"],
  M2_Slope = coef(m_public_2)$unit_id[,"x_year"],
  Growth_Pct_M1 = M1_Slope / M1_Intcpt,
  Growth_Pct_M2 = M2_Slope / M1_Intcpt,
  GPM_Diff = Growth_Pct_M2 - Growth_Pct_M1,
  GPM_Diff_Bucket = case_when(
    GPM_Diff < -0.01 ~ 1,
    GPM_Diff < 0.01 ~ 2,
    GPM_Diff >= 0.01 ~ 3
  ) %>% factor()
) %>%
  pivot_longer(cols=5:6) %>%
  left_join(
    y = public_fyr_normalized %>% distinct(unit_id, carnegie_category, size_category),
    by = "unit_id"
  ) %>%
  ggplot(aes(x=name, y=value)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group=unit_id, color=GPM_Diff_Bucket), alpha = 0.5, linewidth=2) +
  scale_color_brewer(type="div") +
  facet_grid(size_category ~ carnegie_category) +
  theme_classic()
ggsave("../Results/m2_slope_shift_public.png")
```

## Model 2 Estimation: Private

```{r}
m_private_2 <- lmer(formula_2, private_dat3_x5)
summary(m_private_2)
```

```{r}
tibble(
  unit_id = rownames(coef(m_private_1)$unit_id) %>% as.numeric(),
  M1_Intcpt = coef(m_private_1)$unit_id[,"(Intercept)"],
  M1_Slope = coef(m_private_1)$unit_id[,"x_year"],
  M2_Slope = coef(m_private_2)$unit_id[,"x_year"],
  Growth_Pct_M1 = M1_Slope / M1_Intcpt,
  Growth_Pct_M2 = M2_Slope / M1_Intcpt,
  GPM_Diff = Growth_Pct_M2 - Growth_Pct_M1,
  GPM_Diff_Bucket = case_when(
    GPM_Diff < -0.01 ~ 1,
    GPM_Diff < 0.01 ~ 2,
    GPM_Diff >= 0.01 ~ 3
  ) %>% factor()
) %>%
  pivot_longer(cols=5:6) %>%
  left_join(
    y = private_fyr_normalized %>% distinct(unit_id, carnegie_category, size_category),
    by = "unit_id"
  ) %>%
  ggplot(aes(x=name, y=value)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group=unit_id, color=GPM_Diff_Bucket), alpha = 0.5, linewidth=2) +
  scale_color_brewer(type="div") +
  facet_grid(size_category ~ carnegie_category) +
  theme_classic()
ggsave("../Results/m2_slope_shift_private.png")
```

# Model 3: Institutional Actions

```{r}
formula_3 <- formula(
  first_time_enrollment ~
    1
    + x_year
    + x_rural
    + sx_bach # Percent of adults with bachelors degree in the state where the institution is located, standardized
    + sx_hs # The number of high school graduates that year, logged and standardized
    + sx_race_minority
    + sx_pell
    + sx_apps
    + sx_admit_rate
    + sx_price
    # + (1 + x_year | carnegie)
    + (1 + x_year | unit_id)
)
```


# Model Estimation: Public

```{r}
m_public_3 <- lmer(formula = formula_3, data = public_dat3_x5)
summary(m_public_3)
coef(m_public_3)
```

```{r}
m_private_3 <- lmer(formula = formula_3, data = private_dat3_x5)
summary(m_private_3)
coef(m_private_3)
```

The model did not converge. Let's explore what's going on

```{r}
cor(private_dat3_x5$sx_apps, private_dat3_x5$sx_admit_rate) # correlation is ~ -0.5 so not collinear

cor(private_dat3_x5[,c(
  "sx_apps"
  ,"sx_admit_rate"
  ,"sx_pell"
  ,"sx_price"
  ,"sx_race_minority"
  )
  ]
)

glmer(formula = formula_3, data = private_dat3_x5, family = "poisson") # this runs
```

From https://www.geeksforgeeks.org/how-to-resolve-r-glmer-warnings-model-fails-to-converge-and-model-is-nearly-unidentifiable/?ref=asr4

> Model Fails to Converge: This warning suggests that the algorithm used to fit the GLMM did not successfully find a solution that meets convergence criteria. Convergence failure can occure due to various reasons, including:
> * Poor starting values for the optimization algorithm
> * Non-linear relationships between predictors and the response variable
> * High collinearity among predictors
> * Overdispersion or underdispersion in the data
> * Sparse data or small sample sizes

I think we can rule out collinearity despite the natural relationship between total number of applications and admission rate because even then the correlation between them is not close to 1.

I think overdispersion is much more likely given that we were already thinking that might be a problem.

```{r}
glm_public_1 <- glmer(formula = formula_1, data = public_dat3_x5, family = "poisson")
glm_public_2 <- glmer(formula = formula_2, data = public_dat3_x5, family = "poisson")
glm_public_3 <- glmer(formula = formula_3, data = public_dat3_x5, family = "poisson")

glm_private_1 <- glmer(formula = formula_1, data = private_dat3_x5, family = "poisson")
glm_private_2 <- glmer(formula = formula_2, data = private_dat3_x5, family = "poisson")
glm_private_3 <- glmer(formula = formula_3, data = private_dat3_x5, family = "poisson")

```

```{r}
glmer.nb(formula_3, data=public_dat3_x5)
```


```{r}
map(
  .x = list(
    glm_public_1,
    glm_public_2,
    glm_public_3,
    glm_private_1,
    glm_private_2,
    glm_private_3
  ),
  .f = summary
)
```

```{r}
anova(glm_public_1, glm_public_2, glm_public_3)

anova(glm_private_1, glm_private_2, glm_private_3)
```

Let's see if number of iterations does something for the non-convergence

```{r}
lmer(formula = formula_3, data = private_dat3_x5, control = lmerControl())
?lmerControl
```

```{r}
allFit(m_private_3)
```


```{r}
public_dat3_x5 %>%
  select(unit_id, year, pct_race_minority) %>%
  left_join(
    y = public_fyr_normalized,
    by = c("unit_id","year")
  ) %>%
  ggplot(aes(x=year, y=pct_race_minority)) +
  geom_line(aes(group=unit_id, color=growth_bucket)) +
  scale_color_brewer(type = "div")

predict(
  m_public_3, 
  tibble(
    x_year = c(0,4),
    x_rural = c("City/Suburb","City/Suburb"),
    sx_bach = c(0,0),
    sx_hs = c(0,0),
    sx_race_minority = c(-1,1),
    sx_pell = c(0,0),
    sx_apps = c(0,0),
    sx_admit_rate = c(0,0),
    sx_price = c(0,0),
    unit_id = c(200332, 200332)
  )
)

str(m_public_3)
m_public_3@call

count(public_dat3_x5, locale_rural)

public_fyr_normalized %>%
  left_join(
    y = public_dat3_x5 %>%
      filter(year==2019) %>%
      mutate(xvar_cut = cut(pct_race_minority, 3)) %>%
      select(unit_id, xvar_cut),
    by = "unit_id"
  ) %>%
  ggplot(aes(x=x_year, y=normalized_enrollment)) +
  geom_line(mapping=aes(group=unit_id, color=growth_bucket), alpha = 0.9) +
  scale_color_brewer(type="div", palette = "RdYlBu") +
  facet_grid(size_category ~ xvar_cut)

map(
  .x = list(
    "pct_race_minority",
    "pct_pell",
    # "applicants",
    "sx_apps",
    "admissions_rate",
    "avg_net_price_w_aid"
  ),
  .f = function(x) {
    baseline_tertiles <- filter(public_dat3_x5, year==2019)
    baseline_tertiles$xvar_cut <- cut(baseline_tertiles[[x]], 5)
    baseline_tertiles <- select(baseline_tertiles, unit_id, xvar_cut)
    
    public_fyr_normalized %>%
      left_join(
        y = baseline_tertiles,
        by = "unit_id"
      ) %>%
      ggplot(aes(x=x_year, y=normalized_enrollment)) +
      geom_line(mapping=aes(group=unit_id, color=growth_bucket), alpha = 0.9) +
      scale_color_brewer(type="div", palette = "RdYlBu") +
      facet_grid(1 ~ xvar_cut) +
      theme_classic() +
      ggplot2::ggtitle(x)
  }
)
```


# Prototypes?

```{r}
gower_distance <- kmed::distmix(
  data = public_dat3_x5 %>% 
    filter(year==2019) %>%
    select(
      x_rural, sx_bach, sx_hs, 
      sx_race_minority, sx_pell, sx_apps, sx_admit_rate, sx_price
    ),
  method = "gower",
  idnum = 2:8,
  idcat = 1
)

set.seed(20250507)
rkm_result <- kmed::rankkmed(gower_distance, ncluster = 5, m = 10, iterate = 50)

prototypes_baseline <- filter(public_dat3_x5, year==2019)[rkm_result$medoid,]

critics_baseline <- tibble(
  unit_id = filter(public_dat3_x5, year==2019)[["unit_id"]],
  Rank1 = rank(gower_distance[rkm_result$medoid,][1,]),
  Rank2 = rank(gower_distance[rkm_result$medoid,][2,]),
  Rank3 = rank(gower_distance[rkm_result$medoid,][3,]),
  Rank4 = rank(gower_distance[rkm_result$medoid,][4,]),
  Rank5 = rank(gower_distance[rkm_result$medoid,][5,])
) %>%
  mutate(Tot_Rank = Rank1 + Rank2 + Rank3 + Rank4 + Rank5) %>%
  arrange(desc(Tot_Rank)) %>%
  head(5) %>%
  left_join(
    y = public_dat3_x5 %>% filter(year == 2019),
    by = "unit_id"
  )

```

```{r}
gower_b <- kmed::distmix(
  data = public_dat3_x5 %>% 
    filter(year==2019, carnegie=="Baccalaureate") %>%
    select(
      x_rural, sx_bach, sx_hs, 
      sx_race_minority, sx_pell, sx_apps, sx_admit_rate, sx_price
    ),
  method = "gower",
  idnum = 2:8,
  idcat = 1
)

gower_m <- kmed::distmix(
  data = public_dat3_x5 %>% 
    filter(year==2019, carnegie=="M") %>%
    select(
      x_rural, sx_bach, sx_hs, 
      sx_race_minority, sx_pell, sx_apps, sx_admit_rate, sx_price
    ),
  method = "gower",
  idnum = 2:8,
  idcat = 1
)

gower_d <- kmed::distmix(
  data = public_dat3_x5 %>% 
    filter(year==2019, carnegie %in% c("R1","R2")) %>%
    select(
      x_rural, sx_bach, sx_hs, 
      sx_race_minority, sx_pell, sx_apps, sx_admit_rate, sx_price
    ),
  method = "gower",
  idnum = 2:8,
  idcat = 1
)

set.seed(20250507)
kmed_b <- kmed::rankkmed(gower_b, ncluster = 3, m = 10, iterate = 50)
kmed_m <- kmed::rankkmed(gower_m, ncluster = 3, m = 10, iterate = 50)
kmed_d <- kmed::rankkmed(gower_d, ncluster = 3, m = 10, iterate = 50)

prototypes_baseline <- rbind(
  filter(public_dat3_x5, year==2019, carnegie=="Baccalaureate")[kmed_b$medoid,],
  filter(public_dat3_x5, year==2019, carnegie=="M")[kmed_m$medoid,],
  filter(public_dat3_x5, year==2019, carnegie %in% c("R1","R2"))[kmed_d$medoid,]
)

```

```{r}
prototypes_baseline %>%
  select(unit_id) %>%
  left_join(public_dat3_x5, by = "unit_id") %>%
  cross_join(
    y = tibble(
      pred_race_minority = c(-0.5, 0.25, 0, 0.25, 0.5)
    )
  ) %>%
  select(-sx_race_minority) %>%
  rename(sx_race_minority = pred_race_minority) %>%
  mutate(
    pred_fyr = predict(object = m_public_3, newdata=.),
    pred_grp = paste0(unit_id,":",sx_race_minority)
  ) %>%
  group_by(pred_grp) %>%
  mutate(
    pred_fyr_norm = pred_fyr / pred_fyr[which(year==2019)]
  ) %>%
  ggplot(
    aes(x = x_year, y = pred_fyr_norm)
  ) +
  geom_line(aes(group=pred_grp, color = factor(sx_race_minority))) +
  scale_color_brewer(type="div") +
  facet_wrap(~unit_id)

predict(m_public_3, prototypes_baseline)
```




